#Load required library
```{r}
library(edgeR)
library(variancePartition)
library(Seurat)
library(dplyr)
library(tidyverse)
library(data.table)
library(ggplot2)
library(ggrepel)
library(colorRamp2)
library(pheatmap)
library(ggpubr)
library(openxlsx)
```

#1.Read DEGs results from Dream analyis
```{r}
#TH outputs 
TH_Output_PATH <- "/mnt/icbs_shared_storage_general_2022/Benjamaporn/Analysis2026/Analysis202603/Part4_AIDA_Age_Sex_DEGs_Output/Thai_AIDA"
TH_AIDA_Myeloid_pDC_dream_modelFit_eBayes <- readRDS(file.path(TH_Output_PATH, "/TH_AIDA_Myeloid_pDC_dream_modelFit_eBayes.rds"))
TH_AIDA_NK_dream_modelFit_eBayes <- readRDS(file.path(TH_Output_PATH, "/TH_AIDA_NK_dream_modelFit_eBayes.rds"))
TH_AIDA_T_dream_modelFit_eBayes <- readRDS(file.path(TH_Output_PATH, "/TH_AIDA_T_dream_modelFit_eBayes.rds"))
TH_AIDA_B_dream_modelFit_eBayes <- readRDS(file.path(TH_Output_PATH, "/TH_AIDA_B_dream_modelFit_eBayes.rds"))

#SG outputs
SG_Output_PATH <- "/mnt/icbs_shared_storage_general_2022/Benjamaporn/Analysis2026/Analysis202603/Part4_AIDA_Age_Sex_DEGs_Output/SG_AIDA"
SG_AIDA_Myeloid_pDC_dream_modelFit_eBayes <- readRDS(file.path(SG_Output_PATH, "/SG_AIDA_Myeloid_pDC_dream_modelFit_eBayes.rds"))
SG_AIDA_NK_dream_modelFit_eBayes <- readRDS(file.path(SG_Output_PATH, "/SG_AIDA_NK_dream_modelFit_eBayes.rds"))
SG_AIDA_T_dream_modelFit_eBayes <- readRDS(file.path(SG_Output_PATH, "/SG_AIDA_T_dream_modelFit_eBayes.rds"))
SG_AIDA_B_dream_modelFit_eBayes <- readRDS(file.path(SG_Output_PATH, "/SG_AIDA_B_dream_modelFit_eBayes.rds"))

#SG_Malay outputs
SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_eBayes <- readRDS(file.path(SG_Output_PATH, "/SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_eBayes.rds"))
SG_Malay_AIDA_NK_dream_modelFit_eBayes <- readRDS(file.path(SG_Output_PATH, "/SG_Malay_AIDA_NK_dream_modelFit_eBayes.rds"))
SG_Malay_AIDA_T_dream_modelFit_eBayes <- readRDS(file.path(SG_Output_PATH, "/SG_Malay_AIDA_T_dream_modelFit_eBayes.rds"))
SG_Malay_AIDA_B_dream_modelFit_eBayes <- readRDS(file.path(SG_Output_PATH, "/SG_Malay_AIDA_B_dream_modelFit_eBayes.rds"))

#SG_Chinese outputs
SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_eBayes <- readRDS(file.path(SG_Output_PATH, "/SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_eBayes.rds"))
SG_Chinese_AIDA_NK_dream_modelFit_eBayes <- readRDS(file.path(SG_Output_PATH, "/SG_Chinese_AIDA_NK_dream_modelFit_eBayes.rds"))
SG_Chinese_AIDA_T_dream_modelFit_eBayes <- readRDS(file.path(SG_Output_PATH, "/SG_Chinese_AIDA_T_dream_modelFit_eBayes.rds"))
SG_Chinese_AIDA_B_dream_modelFit_eBayes <- readRDS(file.path(SG_Output_PATH, "/SG_Chinese_AIDA_B_dream_modelFit_eBayes.rds"))

#SG_Indian outputs
SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_eBayes <- readRDS(file.path(SG_Output_PATH, "/SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_eBayes.rds"))
SG_Indian_AIDA_NK_dream_modelFit_eBayes <- readRDS(file.path(SG_Output_PATH, "/SG_Indian_AIDA_NK_dream_modelFit_eBayes.rds"))
SG_Indian_AIDA_T_dream_modelFit_eBayes <- readRDS(file.path(SG_Output_PATH, "/SG_Indian_AIDA_T_dream_modelFit_eBayes.rds"))
SG_Indian_AIDA_B_dream_modelFit_eBayes <- readRDS(file.path(SG_Output_PATH, "/SG_Indian_AIDA_B_dream_modelFit_eBayes.rds"))

#KR outputs
KR_Output_PATH <- "/mnt/icbs_shared_storage_general_2022/Benjamaporn/Analysis2026/Analysis202603/Part4_AIDA_Age_Sex_DEGs_Output/KR_AIDA"
KR_AIDA_Myeloid_pDC_dream_modelFit_eBayes <- readRDS(file.path(KR_Output_PATH, "/KR_AIDA_Myeloid_pDC_dream_modelFit_eBayes.rds"))
KR_AIDA_NK_dream_modelFit_eBayes <- readRDS(file.path(KR_Output_PATH, "/KR_AIDA_NK_dream_modelFit_eBayes.rds"))
KR_AIDA_T_dream_modelFit_eBayes <- readRDS(file.path(KR_Output_PATH, "/KR_AIDA_T_dream_modelFit_eBayes.rds"))
KR_AIDA_B_dream_modelFit_eBayes <- readRDS(file.path(KR_Output_PATH, "/KR_AIDA_B_dream_modelFit_eBayes.rds"))

#JP outputs
JP_Output_PATH <- "/mnt/icbs_shared_storage_general_2022/Benjamaporn/Analysis2026/Analysis202603/Part4_AIDA_Age_Sex_DEGs_Output/JP_AIDA"
JP_AIDA_Myeloid_pDC_dream_modelFit_eBayes <- readRDS(file.path(JP_Output_PATH, "/JP_AIDA_Myeloid_pDC_dream_modelFit_eBayes.rds"))
JP_AIDA_NK_dream_modelFit_eBayes <- readRDS(file.path(JP_Output_PATH, "/JP_AIDA_NK_dream_modelFit_eBayes.rds"))
JP_AIDA_T_dream_modelFit_eBayes <- readRDS(file.path(JP_Output_PATH, "/JP_AIDA_T_dream_modelFit_eBayes.rds"))
JP_AIDA_B_dream_modelFit_eBayes <- readRDS(file.path(JP_Output_PATH, "/JP_AIDA_B_dream_modelFit_eBayes.rds"))
```

#Immune Cell types of interest
```{r}
Myeloid_pDC_Author_Annotation_celltype <- c("Myeloid_unknown", "CD14+_Monocyte", "CD16+_Monocyte", "cDC2")
NK_Author_Annotation_celltype <- c("NK_unknown", "CD16+_NK", "CD56+_NK")
T_Author_Annotation_celltype <- c("T_unknown", "CD4+_T_unknown", "CD4+_T_naive", "CD4+_T_cm",  "CD4+_T_cyt", "CD4+_T_em", "CD8+_T_unknown", "CD8+_T_naive", "CD8+_T_GZMBhi", "CD8+_T_GZMKhi", "MAIT",   "gdT_GZMBhi", "gdT_GZMKhi",  "Treg")
B_Author_Annotation_celltype <- c("B_unknown", "naive_B", "memory_B_IGHMlo", "memory_B_IGHMhi")
```

#Thai populations
#2. DEGs counts impacted by Blue ancestry
```{r fig.height=4, fig.width=10}
#No. of DEGs in each cell type
#Count no. of DEGs
#Myeloid_pDC
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref <- list()
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_diffexpressed <- list()
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_sig_count <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
  TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref[[i]] <- topTable(TH_AIDA_Myeloid_pDC_dream_modelFit_eBayes[[i]], coef = "Blue_ref", adjust="BH", number = Inf) 
  TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_diffexpressed[[i]] <- data.frame(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref[[i]]) %>%  
  mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                   logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                   TRUE ~ "Not significant"))
  TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_diffexpressed[[i]]$Gene <- rownames(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_diffexpressed[[i]])
  TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_sig_count[[i]] <- subset(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_diffexpressed[[i]], diffexpressed != "Not significant") %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#NK
TH_AIDA_NK_dream_modelFit_Blue_ref <- list()
TH_AIDA_NK_dream_modelFit_Blue_ref_diffexpressed <- list()
TH_AIDA_NK_dream_modelFit_Blue_ref_sig_count <- list()
for (i in NK_Author_Annotation_celltype) {
  TH_AIDA_NK_dream_modelFit_Blue_ref[[i]] <- topTable(TH_AIDA_NK_dream_modelFit_eBayes[[i]], coef = "Blue_ref", adjust="BH", number = Inf) 
  TH_AIDA_NK_dream_modelFit_Blue_ref_diffexpressed[[i]] <- data.frame(TH_AIDA_NK_dream_modelFit_Blue_ref[[i]]) %>%
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                     logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  TH_AIDA_NK_dream_modelFit_Blue_ref_diffexpressed[[i]]$Gene <- rownames(TH_AIDA_NK_dream_modelFit_Blue_ref_diffexpressed[[i]])
  TH_AIDA_NK_dream_modelFit_Blue_ref_sig_count[[i]] <- subset(TH_AIDA_NK_dream_modelFit_Blue_ref_diffexpressed[[i]], diffexpressed != "Not significant") %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#T
TH_AIDA_T_dream_modelFit_Blue_ref <- list()
TH_AIDA_T_dream_modelFit_Blue_ref_diffexpressed <- list()
TH_AIDA_T_dream_modelFit_Blue_ref_sig_count <- list()
for (i in T_Author_Annotation_celltype) {
  TH_AIDA_T_dream_modelFit_Blue_ref[[i]] <- topTable(TH_AIDA_T_dream_modelFit_eBayes[[i]], coef = "Blue_ref", adjust="BH", number = Inf) 
  TH_AIDA_T_dream_modelFit_Blue_ref_diffexpressed[[i]] <- data.frame(TH_AIDA_T_dream_modelFit_Blue_ref[[i]]) %>%  
  mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                   logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                   TRUE ~ "Not significant"))
  TH_AIDA_T_dream_modelFit_Blue_ref_diffexpressed[[i]]$Gene <- rownames(TH_AIDA_T_dream_modelFit_Blue_ref_diffexpressed[[i]])
  TH_AIDA_T_dream_modelFit_Blue_ref_sig_count[[i]] <- subset(TH_AIDA_T_dream_modelFit_Blue_ref_diffexpressed[[i]], diffexpressed != "Not significant") %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#B
TH_AIDA_B_dream_modelFit_Blue_ref <- list()
TH_AIDA_B_dream_modelFit_Blue_ref_diffexpressed <- list()
TH_AIDA_B_dream_modelFit_Blue_ref_sig_count <- list()
for (i in B_Author_Annotation_celltype) {
  TH_AIDA_B_dream_modelFit_Blue_ref[[i]] <- topTable(TH_AIDA_B_dream_modelFit_eBayes[[i]], coef = "Blue_ref", adjust="BH", number = Inf) 
  TH_AIDA_B_dream_modelFit_Blue_ref_diffexpressed[[i]] <- data.frame(TH_AIDA_B_dream_modelFit_Blue_ref[[i]]) %>%  
  mutate(diffexpressed = case_when(logFC > 1 &   adj.P.Val < 0.05 ~ "Upregulated", 
                                   logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                   TRUE ~ "Not significant"))
  TH_AIDA_B_dream_modelFit_Blue_ref_diffexpressed[[i]]$Gene <- rownames(TH_AIDA_B_dream_modelFit_Blue_ref_diffexpressed[[i]])
  TH_AIDA_B_dream_modelFit_Blue_ref_sig_count[[i]] <- subset(TH_AIDA_B_dream_modelFit_Blue_ref_diffexpressed[[i]], diffexpressed != "Not significant") %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#Obtain only genes with |logFC| > 1 & adj.P.Val < 0.05 by Blue ancestry
TH_DEGs_sig_Blue_ref_counts <- data.frame(
    Author_Annotation_celltype = rep(c("Myeloid_unknown", "CD14+_Monocyte", "CD16+_Monocyte", "cDC2", "NK_unknown", "CD16+_NK", "CD56+_NK", "T_unknown", "CD4+_T_unknown", "CD4+_T_naive", "CD4+_T_cm",  "CD4+_T_cyt", "CD4+_T_em", "CD8+_T_unknown", "CD8+_T_naive", "CD8+_T_GZMBhi", "CD8+_T_GZMKhi", "MAIT", "gdT_GZMBhi", "gdT_GZMKhi",  "Treg", "B_unknown", "naive_B", "memory_B_IGHMlo", "memory_B_IGHMhi"), each=2),
    DEGs=rep(c("Upregulated with Blue ancestry", "Downregulated with Blue ancestry"), 25),
    nGenes=c("0", "0", "15","10", "0", "0", "0", "0", 
             "0", "0", "3", "49", "1", "0", 
             "0", "0", "1", "0", "0", "4", "1", "5", "1", "0", "0", "0", "0", "2", "1", "7", "1", "1", "0", "0", "0", "0", "0", "0", "1", "1", "0", "0",
             "0", "0", "0", "0", "2", "0", "0", "0")
)

#Plotting
TH_DEGs_sig_Blue_ref_counts$Author_Annotation_celltype <- factor(TH_DEGs_sig_Blue_ref_counts$Author_Annotation_celltype, levels = c("Myeloid_unknown", "CD14+_Monocyte", "CD16+_Monocyte", "cDC2", "NK_unknown", "CD16+_NK", "CD56+_NK", "T_unknown", "CD4+_T_unknown", "CD4+_T_naive", "CD4+_T_cm",  "CD4+_T_cyt", "CD4+_T_em", "CD8+_T_unknown", "CD8+_T_naive", "CD8+_T_GZMBhi", "CD8+_T_GZMKhi", "MAIT",   "gdT_GZMBhi", "gdT_GZMKhi",  "Treg", "B_unknown", "naive_B", "memory_B_IGHMlo", "memory_B_IGHMhi"))
TH_DEGs_sig_Blue_ref_counts$nGenes <- as.numeric(as.vector(TH_DEGs_sig_Blue_ref_counts$nGenes))
TH_DEGs_sig_Blue_ref_counts_plot <- ggplot(data=TH_DEGs_sig_Blue_ref_counts, aes(x=Author_Annotation_celltype, y=nGenes, fill=DEGs)) +
  geom_bar(stat="identity", position=position_dodge())+
  geom_text(aes(label=nGenes), vjust=-0.2, color="black", position = position_dodge(0.9), size=3)+ 
  scale_fill_manual(values = c("Upregulated with Blue ancestry" = "#2c828d", "Downregulated with Blue ancestry" ="#8d372c")) +
  theme_minimal() + 
  theme(legend.position = c(0.85, 0.9), legend.title = element_blank())+
  ylab("No. of DEGs") +
  xlab("") +
  ggtitle("Thai at |logFC| > 1 and adj.P.Val < 0.05")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, color = "black"))
```

#3. DEGs counts impacted by Sex 
```{r fig.height=4, fig.width=10}
y_linked_genes_1logFC <- c("RPS4Y1", "DDX3Y", "UTY", "USP9Y", "PRKY", "KDM5D", "EIF1AY", "ZFY", "TTTY14", "LINC00278", "AC244213.1", "RPS4Y2", "AC007244.1", "ZFY-AS1", "NLGN4Y", "TMSB4Y", "AC006157.1", "TMSB4Y", "TTTY10", "TMSB4Y", "TTTY10")

#No. of DEGs in each cell type
#Count no. of DEGs 
#Myeloid_pDC
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale <- list()
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_diffexpressed <- list()
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
  TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale[[i]] <- topTable(TH_AIDA_Myeloid_pDC_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf) 
  TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]] <- data.frame(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant")) 
  TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]])
  TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant")
  TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene)) 
}

#NK
TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale <- list()
TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_diffexpressed <- list()
TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in NK_Author_Annotation_celltype) {
  TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale[[i]] <- topTable(TH_AIDA_NK_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf)
  TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]] <- data.frame(TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]])
  TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant") 
  TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#T
TH_AIDA_T_dream_modelFit_Blue_ref_SexMale <- list()
TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_diffexpressed <- list()
TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in T_Author_Annotation_celltype) {
  TH_AIDA_T_dream_modelFit_Blue_ref_SexMale[[i]] <- topTable(TH_AIDA_T_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf)
  TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]] <- data.frame(TH_AIDA_T_dream_modelFit_Blue_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]])
  TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant") 
  TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#B
TH_AIDA_B_dream_modelFit_Blue_ref_SexMale <- list()
TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_diffexpressed <- list()
TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in B_Author_Annotation_celltype) {
    TH_AIDA_B_dream_modelFit_Blue_ref_SexMale[[i]] <- topTable(TH_AIDA_B_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf)
  TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]] <- data.frame(TH_AIDA_B_dream_modelFit_Blue_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]])
  TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant")
 TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#Checked:
TH_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts <- data.frame(
    Author_Annotation_celltype = rep(c("Myeloid_unknown", "CD14+_Monocyte", "CD16+_Monocyte", "cDC2", "NK_unknown", "CD16+_NK", "CD56+_NK", "T_unknown", "CD4+_T_unknown", "CD4+_T_naive", "CD4+_T_cm",  "CD4+_T_cyt", "CD4+_T_em", "CD8+_T_unknown", "CD8+_T_naive", "CD8+_T_GZMBhi", "CD8+_T_GZMKhi", "MAIT",   "gdT_GZMBhi", "gdT_GZMKhi",  "Treg", "B_unknown", "naive_B", "memory_B_IGHMlo", "memory_B_IGHMhi"), each=2),
    DEGs=rep(c("Upregulated with XY", "Downregulated with XY"), 25),
    nGenes=c("0", "2", "0","4", "1", "8", "0", "2", 
             "0", "1", "7", "3", "0", "1", 
             "1", "2", "0", "2", "3", "4", "0", "3", "0", "2", "0", "2", "3", "17", "1", "3", "2", "2", "0", "2", "0", "3", "0", "3", "0", "3", "0", "2",
             "0", "1", "0", "2", "0", "2", "0", "2")
)

#Plotting
TH_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts$Author_Annotation_celltype <- factor(TH_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts$Author_Annotation_celltype, levels = c("Myeloid_unknown", "CD14+_Monocyte", "CD16+_Monocyte", "cDC2", "NK_unknown", "CD16+_NK", "CD56+_NK", "T_unknown", "CD4+_T_unknown", "CD4+_T_naive", "CD4+_T_cm",  "CD4+_T_cyt", "CD4+_T_em", "CD8+_T_unknown", "CD8+_T_naive", "CD8+_T_GZMBhi", "CD8+_T_GZMKhi", "MAIT", "gdT_GZMBhi", "gdT_GZMKhi",  "Treg", "B_unknown", "naive_B", "memory_B_IGHMlo", "memory_B_IGHMhi"))
TH_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts$nGenes <- as.numeric(as.vector(TH_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts$nGenes))

#Fig. 4C
TH_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts_plot <- ggplot(data=TH_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts, aes(x=Author_Annotation_celltype, y=nGenes, fill=DEGs)) +
  geom_bar(stat="identity", position=position_dodge())+
  geom_text(aes(label=nGenes), vjust=-0.2, color="black", position = position_dodge(0.9), size=3)+ 
  scale_fill_manual(values = c("Upregulated with XY" = "#2986cc", "Downregulated with XY" ="#c90076")) +
  theme_minimal() +
  theme(legend.position = c(0.85, 0.9), legend.title = element_blank())+
  ylab("No. of DEGs") +
  xlab("") +
  ylim(0, 18) +
  ggtitle("Thai at |logFC| > 1 and adj.P.Val < 0.05 (without Y-linked genes)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, color = "black"))
```

```{r}
library(ggpubr)
Gender_col_dict <- c("XY" = "#298ccc", "XX" ="#c90076")
sex_col_gradient <- c("#c90076", "#d33291", "#de66ac", "#e999c8", "#f4cce3", "white", "#d4e6f7", "#a9cee0", "#69aadb","#539ed6",  "#3e92d2",  "#298ccc")
Sex_deg_col_dict <- c("Downregulated"="#c90076", "Not significant"="grey", "Upregulated"="#298ccc")
```

#Sex Differences in CD16+ NK cells in the Thai populations
```{r fig.height=6, fig.width=9}
#CD16+ NK
TH_dge_blue_ref_SexMale_CD16NK <- TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale[["CD16+_NK"]]
TH_dge_blue_ref_SexMale_CD16NK$Gene <- rownames(TH_dge_blue_ref_SexMale_CD16NK)
TH_dge_blue_ref_SexMale_CD16NK_noYlinkedGenes <- subset(TH_dge_blue_ref_SexMale_CD16NK, !(Gene %in% y_linked_genes_1logFC))
TH_dge_blue_ref_SexMale_CD16NK_noYlinkedGenes_diffexpressed_newcutoff <- TH_dge_blue_ref_SexMale_CD16NK_noYlinkedGenes %>%  
  mutate(diffexpressed_0.05FDR = case_when(logFC > 0.5 & adj.P.Val < 0.05 ~ "Upregulated", 
                                   logFC < -0.5 & adj.P.Val < 0.05 ~ "Downregulated",
                                   TRUE ~ "Not significant"))

#Fig. 4D
TH_dge_blue_ref_SexMale_CD16NK_noYlinkedGenes_diffexpressed_newcutoff_plot <- ggplot(TH_dge_blue_ref_SexMale_CD16NK_noYlinkedGenes_diffexpressed_newcutoff, aes(x = logFC, y = -log10(adj.P.Val), color = diffexpressed_0.05FDR)) +
  geom_point(size = 2)  +
  ylab("-log10(adjusted p-value)") + 
  geom_text_repel(aes(label = ifelse(logFC > 0.9 & adj.P.Val < 0.05, Gene, "")), colour = "black", size = 2.5, max.overlaps=Inf, segment.color = 'transparent') + 
  geom_text_repel(aes(label = ifelse(logFC < -0.5 & adj.P.Val < 0.05, Gene, "")), colour = "black", size = 2.5, max.overlaps=Inf, segment.color = 'transparent') + 
  scale_color_manual(values=Sex_deg_col_dict) +
  ggtitle('CD16+NK of XY vs XX among Thai populations without Y-linked genes') + 
  theme_bw() +
  labs(color = '')+ 
  theme(text = element_text(size = 14)) + 
  xlim(-5.6,2)
TH_dge_blue_ref_SexMale_CD16NK_noYlinkedGenes_diffexpressed_newcutoff_plot
```

```{r}
#logFC
#Myeloid_pDC
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_select <- list()
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_DEGs <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale[[i]]$Gene <- rownames(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale[[i]])
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_select[[i]] <- select(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale[[i]], logFC, Gene)
setnames(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_select[[i]], old = "logFC", new = i)
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_DEGs[[i]] <- subset(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], diffexpressed != "Not significant") 
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_DEGs[[i]]$Gene <- rownames(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_DEGs[[i]])
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_DEGs[[i]] <- select(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_DEGs[[i]], logFC, Gene)
setnames(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_DEGs[[i]], old = "logFC", new = i)
}

#NK
TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_select <- list()
TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_DEGs <- list()
for (i in NK_Author_Annotation_celltype) {
TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale[[i]]$Gene <- rownames(TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale[[i]])
TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_select[[i]] <- select(TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale[[i]], logFC, Gene) 
setnames(TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_select[[i]], old = "logFC", new = i)
TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_DEGs[[i]] <- subset(TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], diffexpressed != "Not significant") 
TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_DEGs[[i]]$Gene <- rownames(TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_DEGs[[i]])
TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_DEGs[[i]] <- select(TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_DEGs[[i]], logFC, Gene) 
setnames(TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_DEGs[[i]], old = "logFC", new = i)
}

#T
TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_select <- list()
TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_DEGs <- list()
for (i in T_Author_Annotation_celltype) {
TH_AIDA_T_dream_modelFit_Blue_ref_SexMale[[i]]$Gene <- rownames(TH_AIDA_T_dream_modelFit_Blue_ref_SexMale[[i]])
TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_select[[i]] <- select(TH_AIDA_T_dream_modelFit_Blue_ref_SexMale[[i]], logFC, Gene) 
setnames(TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_select[[i]], old = "logFC", new = i)
TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_DEGs[[i]] <- subset(TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], diffexpressed != "Not significant") 
TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_DEGs[[i]]$Gene <- rownames(TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_DEGs[[i]])
TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_DEGs[[i]] <- select(TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_DEGs[[i]], logFC, Gene) 
setnames(TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_DEGs[[i]], old = "logFC", new = i)
}

#B
TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_select <- list()
TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_DEGs <- list()
for (i in B_Author_Annotation_celltype) {
TH_AIDA_B_dream_modelFit_Blue_ref_SexMale[[i]]$Gene <- rownames(TH_AIDA_B_dream_modelFit_Blue_ref_SexMale[[i]])
TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_select[[i]] <- select(TH_AIDA_B_dream_modelFit_Blue_ref_SexMale[[i]], logFC, Gene) 
setnames(TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_select[[i]], old = "logFC", new = i)
TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_DEGs[[i]] <- subset(TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], diffexpressed != "Not significant") 
TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_DEGs[[i]]$Gene <- rownames(TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_DEGs[[i]])
TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_DEGs[[i]] <- select(TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_DEGs[[i]], logFC, Gene) 
setnames(TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_DEGs[[i]], old = "logFC", new = i)
}

#Merge all cell types to get DEGs list
TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_DEGs <- c(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_DEGs, TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_DEGs, TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_DEGs, TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_DEGs) %>% reduce(full_join, by='Gene') 

#Create matrix
TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes <- c(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_select, TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_select, TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_select, TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_select) %>% reduce(full_join, by='Gene') %>% 
filter(Gene %in% TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_DEGs$Gene)
rownames(TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes) <- TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes$Gene
TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes_mtx <- TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes %>% select(-Gene) %>% replace(is.na(.), 0) %>% as.matrix()
```

```{r}
#adjusted P-value
#Myeloid_pDC
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_p_select <- list()
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_p <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_p_select[[i]] <- select(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale[[i]], adj.P.Val, Gene)
setnames(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_p_select[[i]], old = "adj.P.Val", new = i)
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_p[[i]] <- subset(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], diffexpressed != "Not significant") 
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_p[[i]]$Gene <- rownames(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_p[[i]])
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_p[[i]] <- select(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_p[[i]], adj.P.Val, Gene)
setnames(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_p[[i]], old = "adj.P.Val", new = i)
}

#NK
TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_p_select <- list()
TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_p <- list()
for (i in NK_Author_Annotation_celltype) {
TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_p_select[[i]] <- select(TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale[[i]], adj.P.Val, Gene) 
setnames(TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_p_select[[i]], old = "adj.P.Val", new = i)
TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_p[[i]] <- subset(TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], diffexpressed != "Not significant") 
TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_p[[i]]$Gene <- rownames(TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_p[[i]])
TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_p[[i]] <- select(TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_p[[i]], adj.P.Val, Gene) 
setnames(TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_p[[i]], old = "adj.P.Val", new = i)
}

#T
TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_p_select <- list()
TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_p <- list()
for (i in T_Author_Annotation_celltype) {
TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_p_select[[i]] <- select(TH_AIDA_T_dream_modelFit_Blue_ref_SexMale[[i]], adj.P.Val, Gene) 
setnames(TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_p_select[[i]], old = "adj.P.Val", new = i)
TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_p[[i]] <- subset(TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], diffexpressed != "Not significant") 
TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_p[[i]]$Gene <- rownames(TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_p[[i]])
TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_p[[i]] <- select(TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_p[[i]], adj.P.Val, Gene) 
setnames(TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_p[[i]], old = "adj.P.Val", new = i)
}

#B
TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_p_select <- list()
TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_p <- list()
for (i in B_Author_Annotation_celltype) {
TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_p_select[[i]] <- select(TH_AIDA_B_dream_modelFit_Blue_ref_SexMale[[i]], adj.P.Val, Gene) 
setnames(TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_p_select[[i]], old = "adj.P.Val", new = i)
TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_p[[i]] <- subset(TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], diffexpressed != "Not significant") 
TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_p[[i]]$Gene <- rownames(TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_p[[i]])
TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_p[[i]] <- select(TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_p[[i]], adj.P.Val, Gene) 
setnames(TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_p[[i]], old = "adj.P.Val", new = i)
}

#Merge all cell types to get DEGs list
TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_p <- c(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_p, TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_p, TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_p, TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_p) %>% reduce(full_join, by='Gene') 

#Create matrix
TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_p_df_representative_genes <- c(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_p_select, TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale_p_select, TH_AIDA_T_dream_modelFit_Blue_ref_SexMale_p_select, TH_AIDA_B_dream_modelFit_Blue_ref_SexMale_p_select) %>% reduce(full_join, by='Gene') %>% 
filter(Gene %in% TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_DEGs$Gene)
rownames(TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_p_df_representative_genes) <- TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_p_df_representative_genes$Gene
TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_p_df_representative_genes_mtx <- TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_p_df_representative_genes %>% select(-Gene) %>% replace(is.na(.), 0) %>% as.matrix()
```

```{r fig.height=15, fig.width=10}
library(ComplexHeatmap)
sex_col_gradient <- c("#c90076", "#d33291", "#de66ac", "#e999c8", "#f4cce3", "white", "#d4e6f7", "#a9cee0", "#69aadb","#539ed6",  "#3e92d2",  "#298ccc")
#Fig. 4B
TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes_mtx_plot <- 
Heatmap(TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes_mtx, cluster_rows = TRUE, cluster_columns = FALSE, column_title = "Gene Expression Profiles between XX and XY in Thai AIDA", column_title_gp = gpar(fontsize = 12, fontface = "bold"), heatmap_legend_param = list(title = "logFC", legend_direction = "vertical", title_gp = gpar(fontsize = 12), legend_width = unit(6, "cm"), title_position = "lefttop-rot", legend_height = unit(6, "cm")), border_gp = gpar(col = "black", lty = 1, lwd = 2), col=sex_col_gradient, cell_fun = function(j, i, x, y, w, h, fill) {
    logFC <- TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes_mtx[i, j]
    adj.pval  <- TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_p_df_representative_genes_mtx[i, j]
    if (abs(logFC) > 1 & adj.pval < 0.001) {
        grid.text("***", x, y)
    } else if (abs(logFC) > 1 & adj.pval < 0.01) {
        grid.text("**", x, y)
    }  else if (abs(logFC) > 1 & adj.pval < 0.05) {
       grid.text("*", x, y)
    }
}
)
TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes_mtx_plot
```

#4. DEGs counts impacted by Age 
```{r fig.height=4, fig.width=10}
#No. of DEGs in each cell type
#Count no. of DEGs
#Myeloid_pDC
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_Age <- list()
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_Age_diffexpressed <- list()
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_Age_sig_count <- list()
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_Age_max <- list()
TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_Age_min <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
  TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_Age[[i]] <- topTable(TH_AIDA_Myeloid_pDC_dream_modelFit_eBayes[[i]], coef = "Age", adjust="BH", number = Inf) 
  TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_Age_diffexpressed[[i]] <- data.frame(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_Age[[i]]) %>%  
  mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                   logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                   TRUE ~ "Not significant"))
  TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_Age_diffexpressed[[i]]$Gene <- rownames(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_Age_diffexpressed[[i]])
  TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_Age_sig_count[[i]] <- subset(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_Age_diffexpressed[[i]], diffexpressed != "Not significant") %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
   TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_Age_max[[i]] <- max(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_Age_diffexpressed[[i]]$logFC)
   TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_Age_min[[i]] <- min(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_Age_diffexpressed[[i]]$logFC)
}

#NK
TH_AIDA_NK_dream_modelFit_Blue_ref_Age <- list()
TH_AIDA_NK_dream_modelFit_Blue_ref_Age_diffexpressed <- list()
TH_AIDA_NK_dream_modelFit_Blue_ref_Age_sig_count <- list()
TH_AIDA_NK_dream_modelFit_Blue_ref_Age_max <- list()
TH_AIDA_NK_dream_modelFit_Blue_ref_Age_min <- list()
for (i in NK_Author_Annotation_celltype) {
  TH_AIDA_NK_dream_modelFit_Blue_ref_Age[[i]] <- topTable(TH_AIDA_NK_dream_modelFit_eBayes[[i]], coef = "Age", adjust="BH", number = Inf) 
  TH_AIDA_NK_dream_modelFit_Blue_ref_Age_diffexpressed[[i]] <- data.frame(TH_AIDA_NK_dream_modelFit_Blue_ref_Age[[i]]) %>%
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                     logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  TH_AIDA_NK_dream_modelFit_Blue_ref_Age_diffexpressed[[i]]$Gene <- rownames(TH_AIDA_NK_dream_modelFit_Blue_ref_Age_diffexpressed[[i]])
  TH_AIDA_NK_dream_modelFit_Blue_ref_Age_sig_count[[i]] <- subset(TH_AIDA_NK_dream_modelFit_Blue_ref_Age_diffexpressed[[i]], diffexpressed != "Not significant") %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
  TH_AIDA_NK_dream_modelFit_Blue_ref_Age_max[[i]] <- max(TH_AIDA_NK_dream_modelFit_Blue_ref_Age_diffexpressed[[i]]$logFC)
  TH_AIDA_NK_dream_modelFit_Blue_ref_Age_min[[i]] <- min(TH_AIDA_NK_dream_modelFit_Blue_ref_Age_diffexpressed[[i]]$logFC)
}

#T
TH_AIDA_T_dream_modelFit_Blue_ref_Age <- list()
TH_AIDA_T_dream_modelFit_Blue_ref_Age_diffexpressed <- list()
TH_AIDA_T_dream_modelFit_Blue_ref_Age_sig_count <- list()
TH_AIDA_T_dream_modelFit_Blue_ref_Age_max <- list()
TH_AIDA_T_dream_modelFit_Blue_ref_Age_min <- list()
for (i in T_Author_Annotation_celltype) {
  TH_AIDA_T_dream_modelFit_Blue_ref_Age[[i]] <- topTable(TH_AIDA_T_dream_modelFit_eBayes[[i]], coef = "Age", adjust="BH", number = Inf) 
  TH_AIDA_T_dream_modelFit_Blue_ref_Age_diffexpressed[[i]] <- data.frame(TH_AIDA_T_dream_modelFit_Blue_ref_Age[[i]]) %>%  
  mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                   logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                   TRUE ~ "Not significant"))
  TH_AIDA_T_dream_modelFit_Blue_ref_Age_diffexpressed[[i]]$Gene <- rownames(TH_AIDA_T_dream_modelFit_Blue_ref_Age_diffexpressed[[i]])
  TH_AIDA_T_dream_modelFit_Blue_ref_Age_sig_count[[i]] <- subset(TH_AIDA_T_dream_modelFit_Blue_ref_Age_diffexpressed[[i]], diffexpressed != "Not significant") %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
  TH_AIDA_T_dream_modelFit_Blue_ref_Age_max[[i]] <- max(TH_AIDA_T_dream_modelFit_Blue_ref_Age_diffexpressed[[i]]$logFC)
  TH_AIDA_T_dream_modelFit_Blue_ref_Age_min[[i]] <- min(TH_AIDA_T_dream_modelFit_Blue_ref_Age_diffexpressed[[i]]$logFC)
}

#B
TH_AIDA_B_dream_modelFit_Blue_ref_Age <- list()
TH_AIDA_B_dream_modelFit_Blue_ref_Age_diffexpressed <- list()
TH_AIDA_B_dream_modelFit_Blue_ref_Age_sig_count <- list()
TH_AIDA_B_dream_modelFit_Blue_ref_Age_max <- list()
TH_AIDA_B_dream_modelFit_Blue_ref_Age_min <- list()
for (i in B_Author_Annotation_celltype) {
  TH_AIDA_B_dream_modelFit_Blue_ref_Age[[i]] <- topTable(TH_AIDA_B_dream_modelFit_eBayes[[i]], coef = "Age", adjust="BH", number = Inf) 
  TH_AIDA_B_dream_modelFit_Blue_ref_Age_diffexpressed[[i]] <- data.frame(TH_AIDA_B_dream_modelFit_Blue_ref_Age[[i]]) %>%  
  mutate(diffexpressed = case_when(logFC > 1 &   adj.P.Val < 0.05 ~ "Upregulated", 
                                   logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                   TRUE ~ "Not significant"))
  TH_AIDA_B_dream_modelFit_Blue_ref_Age_diffexpressed[[i]]$Gene <- rownames(TH_AIDA_B_dream_modelFit_Blue_ref_Age_diffexpressed[[i]])
  TH_AIDA_B_dream_modelFit_Blue_ref_Age_sig_count[[i]] <- subset(TH_AIDA_B_dream_modelFit_Blue_ref_Age_diffexpressed[[i]], diffexpressed != "Not significant") %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
  TH_AIDA_B_dream_modelFit_Blue_ref_Age_max[[i]] <- max(TH_AIDA_B_dream_modelFit_Blue_ref_Age_diffexpressed[[i]]$logFC)
  TH_AIDA_B_dream_modelFit_Blue_ref_Age_min[[i]] <- min(TH_AIDA_B_dream_modelFit_Blue_ref_Age_diffexpressed[[i]]$logFC)
}
#ANS: No DEGs pass cut-off by age

#max/min logFC
TH_Age_DE_list <- c(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_Age, TH_AIDA_NK_dream_modelFit_Blue_ref_Age, TH_AIDA_T_dream_modelFit_Blue_ref_Age, TH_AIDA_B_dream_modelFit_Blue_ref_Age)
max(unlist(TH_Age_DE_list$`CD14+_Monocyte`$logFC))
min(unlist(TH_Age_DE_list))
```

#5. Comparing of sex differences with other AIDA populations
#SG_Malay
```{r fig.height=4, fig.width=10}
#No. of DEGs in each cell type
#Count no. of DEGs 
#Myeloid_pDC
SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale <- list()
SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_diffexpressed <- list()
SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
  SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale[[i]] <- topTable(SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf) 
  SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]] <- data.frame(SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant")) 
  SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]])
  SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant")
  SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene)) 
}

#NK
SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale <- list()
SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale_diffexpressed <- list()
SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in NK_Author_Annotation_celltype) {
  SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale[[i]] <- topTable(SG_Malay_AIDA_NK_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf)
  SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]] <- data.frame(SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]])
  SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant") 
  SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#T
SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale <- list()
SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale_diffexpressed <- list()
SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in T_Author_Annotation_celltype) {
  SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale[[i]] <- topTable(SG_Malay_AIDA_T_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf)
  SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]] <- data.frame(SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]])
  SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant") 
  SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#B
SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale <- list()
SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale_diffexpressed <- list()
SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in B_Author_Annotation_celltype) {
    SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale[[i]] <- topTable(SG_Malay_AIDA_B_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf)
  SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]] <- data.frame(SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]])
  SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant")
 SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#Checked:
SG_Malay_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts <- data.frame(
    Author_Annotation_celltype = rep(c("Myeloid_unknown", "CD14+_Monocyte", "CD16+_Monocyte", "cDC2", "NK_unknown", "CD16+_NK", "CD56+_NK", "T_unknown", "CD4+_T_unknown", "CD4+_T_naive", "CD4+_T_cm",  "CD4+_T_cyt", "CD4+_T_em", "CD8+_T_unknown", "CD8+_T_naive", "CD8+_T_GZMBhi", "CD8+_T_GZMKhi", "MAIT",   "gdT_GZMBhi", "gdT_GZMKhi",  "Treg", "B_unknown", "naive_B", "memory_B_IGHMlo", "memory_B_IGHMhi"), each=2),
    DEGs=rep(c("Upregulated with XY", "Downregulated with XY"), 25),
    nGenes=c("0", "2", "3","3", "1", "3", "0", "2", 
             "0", "1", "0", "3", "0", "1", 
             "0", "1", "0", "4", "7", "4", "2", "3", "2", "1", "0", "3", "0", "4", "1", "3", "4", "2", "0", "3", "2", "1", "0", "1", "0", "1", "0", "2",
             "0", "4", "27", "7", "0", "1", "0", "1")
)

#Plotting
SG_Malay_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts$Author_Annotation_celltype <- factor(SG_Malay_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts$Author_Annotation_celltype, levels = c("Myeloid_unknown", "CD14+_Monocyte", "CD16+_Monocyte", "cDC2", "NK_unknown", "CD16+_NK", "CD56+_NK", "T_unknown", "CD4+_T_unknown", "CD4+_T_naive", "CD4+_T_cm",  "CD4+_T_cyt", "CD4+_T_em", "CD8+_T_unknown", "CD8+_T_naive", "CD8+_T_GZMBhi", "CD8+_T_GZMKhi", "MAIT", "gdT_GZMBhi", "gdT_GZMKhi",  "Treg", "B_unknown", "naive_B", "memory_B_IGHMlo", "memory_B_IGHMhi"))
SG_Malay_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts$nGenes <- as.numeric(as.vector(SG_Malay_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts$nGenes))
SG_Malay_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts_plot <- ggplot(data=SG_Malay_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts, aes(x=Author_Annotation_celltype, y=nGenes, fill=DEGs)) +
  geom_bar(stat="identity", position=position_dodge())+
  geom_text(aes(label=nGenes), vjust=-0.2, color="black", position = position_dodge(0.9), size=3)+ 
  scale_fill_manual(values = c("Upregulated with XY" = "#298ccc", "Downregulated with XY" ="#c90076")) +
  theme_minimal() + 
  theme(legend.position = c(0.8, 1), legend.title = element_blank())+
  ylab("No. of DEGs") +
  xlab("")+
  ggtitle("SG_Malay at |logFC| > 1 and adj.P.Val < 0.05 (without Y-linked genes)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, color = "black"))
```
```{r}
#Myeloid_pDC
SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_select <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale[[i]]$Gene <- rownames(SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale[[i]])
SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_select[[i]] <- select(SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale[[i]], logFC, Gene)
setnames(SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#NK
SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale_select <- list()
for (i in NK_Author_Annotation_celltype) {
SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale[[i]]$Gene <- rownames(SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale[[i]])
SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale_select[[i]] <- select(SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale[[i]], logFC, Gene) 
setnames(SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#T
SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale_select <- list()
for (i in T_Author_Annotation_celltype) {
SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale[[i]]$Gene <- rownames(SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale[[i]])
SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale_select[[i]] <- select(SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale[[i]], logFC, Gene) 
setnames(SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#B
SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale_select <- list()
for (i in B_Author_Annotation_celltype) {
SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale[[i]]$Gene <- rownames(SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale[[i]])
SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale_select[[i]] <- select(SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale[[i]], logFC, Gene) 
setnames(SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#Merge all cell types to get DEGs list
SG_Malay_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale <- c(SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale, SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale, SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale, SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale) %>% reduce(full_join, by='Gene') 

#Create matrix and obtain only DEGs by sex from Thai 
SG_Malay_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes <- c(SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_select, SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale_select, SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale_select, SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale_select) %>% reduce(full_join, by='Gene') %>% 
filter(Gene %in% TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_DEGs$Gene)

rownames(SG_Malay_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes) <- SG_Malay_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes$Gene
SG_Malay_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes_mtx <- SG_Malay_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes %>% select(-Gene) %>% replace(is.na(.), 0) %>% as.matrix()
```

#SG_Indian
```{r fig.height=4, fig.width=10}
#No. of DEGs in each cell type
#Count no. of DEGs 
#Myeloid_pDC
SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale <- list()
SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale_diffexpressed <- list()
SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
  SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale[[i]] <- topTable(SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf) 
  SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale_diffexpressed[[i]] <- data.frame(SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant")) 
  SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale_diffexpressed[[i]])
  SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant")
  SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene)) 
}

#NK
SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale <- list()
SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale_diffexpressed <- list()
SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in NK_Author_Annotation_celltype) {
  SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale[[i]] <- topTable(SG_Indian_AIDA_NK_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf)
  SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale_diffexpressed[[i]] <- data.frame(SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale_diffexpressed[[i]])
  SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant") 
  SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#T
SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale <- list()
SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale_diffexpressed <- list()
SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in T_Author_Annotation_celltype) {
  SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale[[i]] <- topTable(SG_Indian_AIDA_T_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf)
  SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale_diffexpressed[[i]] <- data.frame(SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale_diffexpressed[[i]])
  SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant") 
  SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#B
SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale <- list()
SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale_diffexpressed <- list()
SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in B_Author_Annotation_celltype) {
    SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale[[i]] <- topTable(SG_Indian_AIDA_B_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf)
  SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale_diffexpressed[[i]] <- data.frame(SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale_diffexpressed[[i]])
  SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant")
 SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#Checked:
SG_Indian_DEGs_sig_Orange_ref_SexMale_noYlinkedGenes_counts <- data.frame(
    Author_Annotation_celltype = rep(c("Myeloid_unknown", "CD14+_Monocyte", "CD16+_Monocyte", "cDC2", "NK_unknown", "CD16+_NK", "CD56+_NK", "T_unknown", "CD4+_T_unknown", "CD4+_T_naive", "CD4+_T_cm",  "CD4+_T_cyt", "CD4+_T_em", "CD8+_T_unknown", "CD8+_T_naive", "CD8+_T_GZMBhi", "CD8+_T_GZMKhi", "MAIT",   "gdT_GZMBhi", "gdT_GZMKhi",  "Treg", "B_unknown", "naive_B", "memory_B_IGHMlo", "memory_B_IGHMhi"), each=2),
    DEGs=rep(c("Upregulated with XY", "Downregulated with XY"), 25),
    nGenes=c("0", "3", "3","7", "0", "4", "0", "64", 
             "0", "1", "2", "4", "0", "1", 
             "0", "1", "1", "4", "14", "11", "2", "3", "0", "2", "1", "3", "2", "4", "5", "4", "4", "3", "0", "3", "0", "2", "0", "2", "0", "1", "2", "7",
             "0", "1", "2225", "20", "1", "1", "0", "2")
)

#Plotting
SG_Indian_DEGs_sig_Orange_ref_SexMale_noYlinkedGenes_counts$Author_Annotation_celltype <- factor(SG_Indian_DEGs_sig_Orange_ref_SexMale_noYlinkedGenes_counts$Author_Annotation_celltype, levels = c("Myeloid_unknown", "CD14+_Monocyte", "CD16+_Monocyte", "cDC2", "NK_unknown", "CD16+_NK", "CD56+_NK", "T_unknown", "CD4+_T_unknown", "CD4+_T_naive", "CD4+_T_cm",  "CD4+_T_cyt", "CD4+_T_em", "CD8+_T_unknown", "CD8+_T_naive", "CD8+_T_GZMBhi", "CD8+_T_GZMKhi", "MAIT", "gdT_GZMBhi", "gdT_GZMKhi",  "Treg", "B_unknown", "naive_B", "memory_B_IGHMlo", "memory_B_IGHMhi"))
SG_Indian_DEGs_sig_Orange_ref_SexMale_noYlinkedGenes_counts$nGenes <- as.numeric(as.vector(SG_Indian_DEGs_sig_Orange_ref_SexMale_noYlinkedGenes_counts$nGenes))
SG_Indian_DEGs_sig_Orange_ref_SexMale_noYlinkedGenes_counts_plot <- ggplot(data=SG_Indian_DEGs_sig_Orange_ref_SexMale_noYlinkedGenes_counts, aes(x=Author_Annotation_celltype, y=nGenes, fill=DEGs)) +
  geom_bar(stat="identity", position=position_dodge())+
  geom_text(aes(label=nGenes), vjust=-0.2, color="black", position = position_dodge(0.9), size=3)+ 
  scale_fill_manual(values = c("Upregulated with XY" = "#298ccc", "Downregulated with XY" ="#c90076")) +
  theme_minimal() + 
  theme(legend.position = c(0.80, 1), legend.title = element_blank())+
  ylab("No. of DEGs") +
  xlab("")+
  ggtitle("SG_Indian at |logFC| > 1 and adj.P.Val < 0.05 (without Y-linked genes)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, color = "black"))
```
```{r}
#Myeloid_pDC
SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale_select <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale[[i]]$Gene <- rownames(SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale[[i]])
SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale_select[[i]] <- select(SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale[[i]], logFC, Gene)
setnames(SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#NK
SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale_select <- list()
for (i in NK_Author_Annotation_celltype) {
SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale[[i]]$Gene <- rownames(SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale[[i]])
SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale_select[[i]] <- select(SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale[[i]], logFC, Gene) 
setnames(SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#T
SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale_select <- list()
for (i in T_Author_Annotation_celltype) {
SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale[[i]]$Gene <- rownames(SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale[[i]])
SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale_select[[i]] <- select(SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale[[i]], logFC, Gene) 
setnames(SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#B
SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale_select <- list()
for (i in B_Author_Annotation_celltype) {
SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale[[i]]$Gene <- rownames(SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale[[i]])
SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale_select[[i]] <- select(SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale[[i]], logFC, Gene) 
setnames(SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#Merge all cell types to get DEGs list
SG_Indian_AIDA_Allcelltypes_dream_modelFit_Orange_ref_SexMale <- c(SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale, SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale, SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale, SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale) %>% reduce(full_join, by='Gene') 

#Create matrix and obtain only DEGs by sex from Thai 
SG_Indian_AIDA_Allcelltypes_dream_modelFit_Orange_ref_SexMale_df_representative_genes <- c(SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale_select, SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale_select, SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale_select, SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale_select) %>% reduce(full_join, by='Gene') %>% 
filter(Gene %in% TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_DEGs$Gene)

rownames(SG_Indian_AIDA_Allcelltypes_dream_modelFit_Orange_ref_SexMale_df_representative_genes) <- SG_Indian_AIDA_Allcelltypes_dream_modelFit_Orange_ref_SexMale_df_representative_genes$Gene
SG_Indian_AIDA_Allcelltypes_dream_modelFit_Orange_ref_SexMale_df_representative_genes_mtx <- SG_Indian_AIDA_Allcelltypes_dream_modelFit_Orange_ref_SexMale_df_representative_genes %>% select(-Gene) %>% replace(is.na(.), 0) %>% as.matrix()
```

#SG_Chinese
```{r fig.height=4, fig.width=10}
#No. of DEGs in each cell type
#Count no. of DEGs 
#Myeloid_pDC
SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale <- list()
SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_diffexpressed <- list()
SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
  SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale[[i]] <- topTable(SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf) 
  SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]] <- data.frame(SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant")) 
  SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]])
  SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant")
  SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene)) 
}

#NK
SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale <- list()
SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale_diffexpressed <- list()
SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in NK_Author_Annotation_celltype) {
  SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale[[i]] <- topTable(SG_Chinese_AIDA_NK_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf)
  SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]] <- data.frame(SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]])
  SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant") 
  SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#T
SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale <- list()
SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale_diffexpressed <- list()
SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in T_Author_Annotation_celltype) {
  SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale[[i]] <- topTable(SG_Chinese_AIDA_T_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf)
  SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]] <- data.frame(SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]])
  SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant") 
  SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#B
SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale <- list()
SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale_diffexpressed <- list()
SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in B_Author_Annotation_celltype) {
    SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale[[i]] <- topTable(SG_Chinese_AIDA_B_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf)
  SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]] <- data.frame(SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]])
  SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant")
 SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#Checked:
SG_Chinese_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts <- data.frame(
    Author_Annotation_celltype = rep(c("Myeloid_unknown", "CD14+_Monocyte", "CD16+_Monocyte", "cDC2", "NK_unknown", "CD16+_NK", "CD56+_NK", "T_unknown", "CD4+_T_unknown", "CD4+_T_naive", "CD4+_T_cm",  "CD4+_T_cyt", "CD4+_T_em", "CD8+_T_unknown", "CD8+_T_naive", "CD8+_T_GZMBhi", "CD8+_T_GZMKhi", "MAIT",   "gdT_GZMBhi", "gdT_GZMKhi",  "Treg", "B_unknown", "naive_B", "memory_B_IGHMlo", "memory_B_IGHMhi"), each=2),
    DEGs=rep(c("Upregulated with XY", "Downregulated with XY"), 25),
    nGenes=c("0", "2", "0","2", "0", "2", "0", "2", 
             "0", "1", "3", "10", "0", "1", 
             "0", "1", "0", "3", "0", "5", "0", "3", "0", "1", "0", "2", "0", "3", "1", "3", "1", "2", "0", "3", "0", "2", "0", "2", "0", "2", "0", "2",
             "0", "6", "0", "12", "0", "1", "0", "1")
)

#Plotting
SG_Chinese_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts$Author_Annotation_celltype <- factor(SG_Chinese_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts$Author_Annotation_celltype, levels = c("Myeloid_unknown", "CD14+_Monocyte", "CD16+_Monocyte", "cDC2", "NK_unknown", "CD16+_NK", "CD56+_NK", "T_unknown", "CD4+_T_unknown", "CD4+_T_naive", "CD4+_T_cm",  "CD4+_T_cyt", "CD4+_T_em", "CD8+_T_unknown", "CD8+_T_naive", "CD8+_T_GZMBhi", "CD8+_T_GZMKhi", "MAIT", "gdT_GZMBhi", "gdT_GZMKhi",  "Treg", "B_unknown", "naive_B", "memory_B_IGHMlo", "memory_B_IGHMhi"))
SG_Chinese_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts$nGenes <- as.numeric(as.vector(SG_Chinese_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts$nGenes))
SG_Chinese_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts_plot <- ggplot(data=SG_Chinese_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts, aes(x=Author_Annotation_celltype, y=nGenes, fill=DEGs)) +
  geom_bar(stat="identity", position=position_dodge())+
  geom_text(aes(label=nGenes), vjust=-0.2, color="black", position = position_dodge(0.9), size=3)+ 
  scale_fill_manual(values = c("Upregulated with XY" = "#298ccc", "Downregulated with XY" ="#c90076")) +
  theme_minimal() + 
  theme(legend.position = c(0.80, 1), legend.title = element_blank())+
  ylab("No. of DEGs") +
  xlab("") +
  ylim(0, 26) +
  ggtitle("SG_Chinese at |logFC| > 1 and adj.P.Val < 0.05 (without Y-linked genes)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, color = "black"))
```
```{r}
#Myeloid_pDC
SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_select <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale[[i]]$Gene <- rownames(SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale[[i]])
SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_select[[i]] <- select(SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale[[i]], logFC, Gene)
setnames(SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#NK
SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale_select <- list()
for (i in NK_Author_Annotation_celltype) {
SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale[[i]]$Gene <- rownames(SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale[[i]])
SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale_select[[i]] <- select(SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale[[i]], logFC, Gene) 
setnames(SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#T
SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale_select <- list()
for (i in T_Author_Annotation_celltype) {
SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale[[i]]$Gene <- rownames(SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale[[i]])
SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale_select[[i]] <- select(SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale[[i]], logFC, Gene) 
setnames(SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#B
SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale_select <- list()
for (i in B_Author_Annotation_celltype) {
SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale[[i]]$Gene <- rownames(SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale[[i]])
SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale_select[[i]] <- select(SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale[[i]], logFC, Gene) 
setnames(SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#Merge all cell types to get DEGs list
SG_Chinese_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale <- c(SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale, SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale, SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale, SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale) %>% reduce(full_join, by='Gene') 

#Create matrix and obtain only DEGs by sex from Thai 
SG_Chinese_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes <- c(SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale_select, SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale_select, SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale_select, SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale_select) %>% reduce(full_join, by='Gene') %>% 
filter(Gene %in% TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_DEGs$Gene)

rownames(SG_Chinese_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes) <- SG_Chinese_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes$Gene
SG_Chinese_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes_mtx <- SG_Chinese_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes %>% select(-Gene) %>% replace(is.na(.), 0) %>% as.matrix()
```

#Korean
```{r fig.height=4, fig.width=10}
#No. of DEGs in each cell type
#Count no. of DEGs 
#Myeloid_pDC
KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale <- list()
KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_diffexpressed <- list()
KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
  KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale[[i]] <- topTable(KR_AIDA_Myeloid_pDC_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf) 
  KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]] <- data.frame(KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant")) 
  KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]])
  KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant")
  KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene)) 
}

#NK
KR_AIDA_NK_dream_modelFit_Red_ref_SexMale <- list()
KR_AIDA_NK_dream_modelFit_Red_ref_SexMale_diffexpressed <- list()
KR_AIDA_NK_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
KR_AIDA_NK_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in NK_Author_Annotation_celltype) {
  KR_AIDA_NK_dream_modelFit_Red_ref_SexMale[[i]] <- topTable(KR_AIDA_NK_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf)
  KR_AIDA_NK_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]] <- data.frame(KR_AIDA_NK_dream_modelFit_Red_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  KR_AIDA_NK_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(KR_AIDA_NK_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]])
  KR_AIDA_NK_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(KR_AIDA_NK_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant") 
  KR_AIDA_NK_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- KR_AIDA_NK_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#T
KR_AIDA_T_dream_modelFit_Red_ref_SexMale <- list()
KR_AIDA_T_dream_modelFit_Red_ref_SexMale_diffexpressed <- list()
KR_AIDA_T_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
KR_AIDA_T_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in T_Author_Annotation_celltype) {
  KR_AIDA_T_dream_modelFit_Red_ref_SexMale[[i]] <- topTable(KR_AIDA_T_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf)
  KR_AIDA_T_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]] <- data.frame(KR_AIDA_T_dream_modelFit_Red_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  KR_AIDA_T_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(KR_AIDA_T_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]])
  KR_AIDA_T_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(KR_AIDA_T_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant") 
  KR_AIDA_T_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- KR_AIDA_T_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#B
KR_AIDA_B_dream_modelFit_Red_ref_SexMale <- list()
KR_AIDA_B_dream_modelFit_Red_ref_SexMale_diffexpressed <- list()
KR_AIDA_B_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
KR_AIDA_B_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in B_Author_Annotation_celltype) {
    KR_AIDA_B_dream_modelFit_Red_ref_SexMale[[i]] <- topTable(KR_AIDA_B_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf)
  KR_AIDA_B_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]] <- data.frame(KR_AIDA_B_dream_modelFit_Red_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  KR_AIDA_B_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(KR_AIDA_B_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]])
  KR_AIDA_B_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(KR_AIDA_B_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant")
 KR_AIDA_B_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- KR_AIDA_B_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#Checked:
KR_DEGs_sig_Red_ref_SexMale_noYlinkedGenes_counts <- data.frame(
    Author_Annotation_celltype = rep(c("Myeloid_unknown", "CD14+_Monocyte", "CD16+_Monocyte", "cDC2", "NK_unknown", "CD16+_NK", "CD56+_NK", "T_unknown", "CD4+_T_unknown", "CD4+_T_naive", "CD4+_T_cm",  "CD4+_T_cyt", "CD4+_T_em", "CD8+_T_unknown", "CD8+_T_naive", "CD8+_T_GZMBhi", "CD8+_T_GZMKhi", "MAIT",   "gdT_GZMBhi", "gdT_GZMKhi",  "Treg", "B_unknown", "naive_B", "memory_B_IGHMlo", "memory_B_IGHMhi"), each=2),
    DEGs=rep(c("Upregulated with XY", "Downregulated with XY"), 25),
    nGenes=c("0", "2", "1","2", "0", "2", "0", "2", 
             "1", "2", "2", "4", "0", "1", 
             "0", "3", "0", "3", "2", "3", "0", "3", "0", "1", "0", "3", "0", "3", "1", "3", "0", "2", "0", "3", "0", "3", "0", "1", "0", "1", "0", "2",
             "0", "1", "0", "2", "0", "4", "0", "2")
)

#Plotting
KR_DEGs_sig_Red_ref_SexMale_noYlinkedGenes_counts$Author_Annotation_celltype <- factor(KR_DEGs_sig_Red_ref_SexMale_noYlinkedGenes_counts$Author_Annotation_celltype, levels = c("Myeloid_unknown", "CD14+_Monocyte", "CD16+_Monocyte", "cDC2", "NK_unknown", "CD16+_NK", "CD56+_NK", "T_unknown", "CD4+_T_unknown", "CD4+_T_naive", "CD4+_T_cm",  "CD4+_T_cyt", "CD4+_T_em", "CD8+_T_unknown", "CD8+_T_naive", "CD8+_T_GZMBhi", "CD8+_T_GZMKhi", "MAIT", "gdT_GZMBhi", "gdT_GZMKhi",  "Treg", "B_unknown", "naive_B", "memory_B_IGHMlo", "memory_B_IGHMhi"))
KR_DEGs_sig_Red_ref_SexMale_noYlinkedGenes_counts$nGenes <- as.numeric(as.vector(KR_DEGs_sig_Red_ref_SexMale_noYlinkedGenes_counts$nGenes))
KR_DEGs_sig_Red_ref_SexMale_noYlinkedGenes_counts_plot <- ggplot(data=KR_DEGs_sig_Red_ref_SexMale_noYlinkedGenes_counts, aes(x=Author_Annotation_celltype, y=nGenes, fill=DEGs)) +
  geom_bar(stat="identity", position=position_dodge())+
  geom_text(aes(label=nGenes), vjust=-0.2, color="black", position = position_dodge(0.9), size=3)+ 
  scale_fill_manual(values = c("Upregulated with XY" = "#298ccc", "Downregulated with XY" ="#c90076")) +
  theme_minimal() + 
  theme(legend.position = c(0.80, 1), legend.title = element_blank())+
  ylab("No. of DEGs") +
  xlab("") +
  ylim(0, 26) +
  ggtitle("Korean at |logFC| > 1 and adj.P.Val < 0.05 (without Y-linked genes)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, color = "black"))
```
```{r}
#Myeloid_pDC
KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_select <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale[[i]]$Gene <- rownames(KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale[[i]])
KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_select[[i]] <- select(KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale[[i]], logFC, Gene)
setnames(KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#NK
KR_AIDA_NK_dream_modelFit_Red_ref_SexMale_select <- list()
for (i in NK_Author_Annotation_celltype) {
KR_AIDA_NK_dream_modelFit_Red_ref_SexMale[[i]]$Gene <- rownames(KR_AIDA_NK_dream_modelFit_Red_ref_SexMale[[i]])
KR_AIDA_NK_dream_modelFit_Red_ref_SexMale_select[[i]] <- select(KR_AIDA_NK_dream_modelFit_Red_ref_SexMale[[i]], logFC, Gene) 
setnames(KR_AIDA_NK_dream_modelFit_Red_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#T
KR_AIDA_T_dream_modelFit_Red_ref_SexMale_select <- list()
for (i in T_Author_Annotation_celltype) {
KR_AIDA_T_dream_modelFit_Red_ref_SexMale[[i]]$Gene <- rownames(KR_AIDA_T_dream_modelFit_Red_ref_SexMale[[i]])
KR_AIDA_T_dream_modelFit_Red_ref_SexMale_select[[i]] <- select(KR_AIDA_T_dream_modelFit_Red_ref_SexMale[[i]], logFC, Gene) 
setnames(KR_AIDA_T_dream_modelFit_Red_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#B
KR_AIDA_B_dream_modelFit_Red_ref_SexMale_select <- list()
for (i in B_Author_Annotation_celltype) {
KR_AIDA_B_dream_modelFit_Red_ref_SexMale[[i]]$Gene <- rownames(KR_AIDA_B_dream_modelFit_Red_ref_SexMale[[i]])
KR_AIDA_B_dream_modelFit_Red_ref_SexMale_select[[i]] <- select(KR_AIDA_B_dream_modelFit_Red_ref_SexMale[[i]], logFC, Gene) 
setnames(KR_AIDA_B_dream_modelFit_Red_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#Merge all cell types to get DEGs list
KR_AIDA_Allcelltypes_dream_modelFit_Red_ref_SexMale <- c(KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale, KR_AIDA_NK_dream_modelFit_Red_ref_SexMale, KR_AIDA_T_dream_modelFit_Red_ref_SexMale, KR_AIDA_B_dream_modelFit_Red_ref_SexMale) %>% reduce(full_join, by='Gene') 

#Create matrix and obtain only DEGs by sex from Thai 
KR_AIDA_Allcelltypes_dream_modelFit_Red_ref_SexMale_df_representative_genes <- c(KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_select, KR_AIDA_NK_dream_modelFit_Red_ref_SexMale_select, KR_AIDA_T_dream_modelFit_Red_ref_SexMale_select, KR_AIDA_B_dream_modelFit_Red_ref_SexMale_select) %>% reduce(full_join, by='Gene') %>% filter(Gene %in% TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_DEGs$Gene)

rownames(KR_AIDA_Allcelltypes_dream_modelFit_Red_ref_SexMale_df_representative_genes) <- KR_AIDA_Allcelltypes_dream_modelFit_Red_ref_SexMale_df_representative_genes$Gene
KR_AIDA_Allcelltypes_dream_modelFit_Red_ref_SexMale_df_representative_genes_mtx <- KR_AIDA_Allcelltypes_dream_modelFit_Red_ref_SexMale_df_representative_genes %>% select(-Gene) %>% replace(is.na(.), 0) %>% as.matrix()
```

#Japanese
```{r fig.height=4, fig.width=10}
#No. of DEGs in each cell type
#Count no. of DEGs 
#Myeloid_pDC
JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale <- list()
JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_diffexpressed <- list()
JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
  JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale[[i]] <- topTable(JP_AIDA_Myeloid_pDC_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf) 
  JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]] <- data.frame(JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant")) 
  JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]])
  JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant")
  JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene)) 
}

#NK
JP_AIDA_NK_dream_modelFit_Red_ref_SexMale <- list()
JP_AIDA_NK_dream_modelFit_Red_ref_SexMale_diffexpressed <- list()
JP_AIDA_NK_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
JP_AIDA_NK_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in NK_Author_Annotation_celltype) {
  JP_AIDA_NK_dream_modelFit_Red_ref_SexMale[[i]] <- topTable(JP_AIDA_NK_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf)
  JP_AIDA_NK_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]] <- data.frame(JP_AIDA_NK_dream_modelFit_Red_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  JP_AIDA_NK_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(JP_AIDA_NK_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]])
  JP_AIDA_NK_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(JP_AIDA_NK_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant") 
  JP_AIDA_NK_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- JP_AIDA_NK_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#T
JP_AIDA_T_dream_modelFit_Red_ref_SexMale <- list()
JP_AIDA_T_dream_modelFit_Red_ref_SexMale_diffexpressed <- list()
JP_AIDA_T_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
JP_AIDA_T_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in T_Author_Annotation_celltype) {
  JP_AIDA_T_dream_modelFit_Red_ref_SexMale[[i]] <- topTable(JP_AIDA_T_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf)
  JP_AIDA_T_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]] <- data.frame(JP_AIDA_T_dream_modelFit_Red_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  JP_AIDA_T_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(JP_AIDA_T_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]])
  JP_AIDA_T_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(JP_AIDA_T_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant") 
  JP_AIDA_T_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- JP_AIDA_T_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#B
JP_AIDA_B_dream_modelFit_Red_ref_SexMale <- list()
JP_AIDA_B_dream_modelFit_Red_ref_SexMale_diffexpressed <- list()
JP_AIDA_B_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed <- list()
JP_AIDA_B_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_sig_count <- list()
for (i in B_Author_Annotation_celltype) {
    JP_AIDA_B_dream_modelFit_Red_ref_SexMale[[i]] <- topTable(JP_AIDA_B_dream_modelFit_eBayes[[i]], coef = "SexMale", adjust="BH", number = Inf)
  JP_AIDA_B_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]] <- data.frame(JP_AIDA_B_dream_modelFit_Red_ref_SexMale[[i]]) %>%  
    mutate(diffexpressed = case_when(logFC > 1 & adj.P.Val < 0.05 ~ "Upregulated", 
                                logFC < -1 & adj.P.Val < 0.05 ~ "Downregulated",
                                TRUE ~ "Not significant"))
  JP_AIDA_B_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]]$Gene <- rownames(JP_AIDA_B_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]])
  JP_AIDA_B_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] <- subset(JP_AIDA_B_dream_modelFit_Red_ref_SexMale_diffexpressed[[i]], !(Gene %in% y_linked_genes_1logFC) & diffexpressed != "Not significant")
 JP_AIDA_B_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_sig_count[[i]] <- JP_AIDA_B_dream_modelFit_Red_ref_SexMale_noYlinkedGenes_diffexpressed[[i]] %>% group_by(diffexpressed) %>% summarise(DEGs = n_distinct(Gene))
}

#Checked:
JP_DEGs_sig_Red_ref_SexMale_noYlinkedGenes_counts <- data.frame(
    Author_Annotation_celltype = rep(c("Myeloid_unknown", "CD14+_Monocyte", "CD16+_Monocyte", "cDC2", "NK_unknown", "CD16+_NK", "CD56+_NK", "T_unknown", "CD4+_T_unknown", "CD4+_T_naive", "CD4+_T_cm",  "CD4+_T_cyt", "CD4+_T_em", "CD8+_T_unknown", "CD8+_T_naive", "CD8+_T_GZMBhi", "CD8+_T_GZMKhi", "MAIT",   "gdT_GZMBhi", "gdT_GZMKhi",  "Treg", "B_unknown", "naive_B", "memory_B_IGHMlo", "memory_B_IGHMhi"), each=2),
    DEGs=rep(c("Upregulated with XY", "Downregulated with XY"), 25),
    nGenes=c("0", "2", "1","3", "0", "2", "0", "2", 
             "1", "1", "4", "3", "0", "1", 
             "0", "1", "0", "4", "0", "4", "0", "3", "0", "1", "0", "2", "3", "3", "0", "3", "0", "2", "0", "1", "0", "2", "0", "1", "0", "1", "0", "2",
             "0", "1", "0", "1", "0", "1", "0", "1")
)

#Plotting
JP_DEGs_sig_Red_ref_SexMale_noYlinkedGenes_counts$Author_Annotation_celltype <- factor(JP_DEGs_sig_Red_ref_SexMale_noYlinkedGenes_counts$Author_Annotation_celltype, levels = c("Myeloid_unknown", "CD14+_Monocyte", "CD16+_Monocyte", "cDC2", "NK_unknown", "CD16+_NK", "CD56+_NK", "T_unknown", "CD4+_T_unknown", "CD4+_T_naive", "CD4+_T_cm",  "CD4+_T_cyt", "CD4+_T_em", "CD8+_T_unknown", "CD8+_T_naive", "CD8+_T_GZMBhi", "CD8+_T_GZMKhi", "MAIT", "gdT_GZMBhi", "gdT_GZMKhi",  "Treg", "B_unknown", "naive_B", "memory_B_IGHMlo", "memory_B_IGHMhi"))
JP_DEGs_sig_Red_ref_SexMale_noYlinkedGenes_counts$nGenes <- as.numeric(as.vector(JP_DEGs_sig_Red_ref_SexMale_noYlinkedGenes_counts$nGenes))
JP_DEGs_sig_Red_ref_SexMale_noYlinkedGenes_counts_plot <- ggplot(data=JP_DEGs_sig_Red_ref_SexMale_noYlinkedGenes_counts, aes(x=Author_Annotation_celltype, y=nGenes, fill=DEGs)) +
  geom_bar(stat="identity", position=position_dodge())+
  geom_text(aes(label=nGenes), vjust=-0.2, color="black", position = position_dodge(0.9), size=3)+ 
  scale_fill_manual(values = c("Upregulated with XY" = "#298ccc", "Downregulated with XY" ="#c90076")) +
  theme_minimal() + 
  theme(legend.position = c(0.80, 1), legend.title = element_blank())+
  ylab("No. of DEGs") +
  xlab("") +
  ylim(0, 26) +
  ggtitle("Japanese at |logFC| > 1 and adj.P.Val < 0.05 (without Y-linked genes)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, color = "black"))
```

```{r}
#Myeloid_pDC
JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_select <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale[[i]]$Gene <- rownames(JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale[[i]])
JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_select[[i]] <- select(JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale[[i]], logFC, Gene)
setnames(JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#NK
JP_AIDA_NK_dream_modelFit_Red_ref_SexMale_select <- list()
for (i in NK_Author_Annotation_celltype) {
JP_AIDA_NK_dream_modelFit_Red_ref_SexMale[[i]]$Gene <- rownames(JP_AIDA_NK_dream_modelFit_Red_ref_SexMale[[i]])
JP_AIDA_NK_dream_modelFit_Red_ref_SexMale_select[[i]] <- select(JP_AIDA_NK_dream_modelFit_Red_ref_SexMale[[i]], logFC, Gene) 
setnames(JP_AIDA_NK_dream_modelFit_Red_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#T
JP_AIDA_T_dream_modelFit_Red_ref_SexMale_select <- list()
for (i in T_Author_Annotation_celltype) {
JP_AIDA_T_dream_modelFit_Red_ref_SexMale[[i]]$Gene <- rownames(JP_AIDA_T_dream_modelFit_Red_ref_SexMale[[i]])
JP_AIDA_T_dream_modelFit_Red_ref_SexMale_select[[i]] <- select(JP_AIDA_T_dream_modelFit_Red_ref_SexMale[[i]], logFC, Gene) 
setnames(JP_AIDA_T_dream_modelFit_Red_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#B
JP_AIDA_B_dream_modelFit_Red_ref_SexMale_select <- list()
for (i in B_Author_Annotation_celltype) {
JP_AIDA_B_dream_modelFit_Red_ref_SexMale[[i]]$Gene <- rownames(JP_AIDA_B_dream_modelFit_Red_ref_SexMale[[i]])
JP_AIDA_B_dream_modelFit_Red_ref_SexMale_select[[i]] <- select(JP_AIDA_B_dream_modelFit_Red_ref_SexMale[[i]], logFC, Gene) 
setnames(JP_AIDA_B_dream_modelFit_Red_ref_SexMale_select[[i]], old = "logFC", new = i)
}

#Merge all cell types to get DEGs list
JP_AIDA_Allcelltypes_dream_modelFit_Red_ref_SexMale <- c(JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale, JP_AIDA_NK_dream_modelFit_Red_ref_SexMale, JP_AIDA_T_dream_modelFit_Red_ref_SexMale, JP_AIDA_B_dream_modelFit_Red_ref_SexMale) %>% reduce(full_join, by='Gene') 

#Create matrix and obtain only DEGs by sex from Thai 
JP_AIDA_Allcelltypes_dream_modelFit_Red_ref_SexMale_df_representative_genes <- c(JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale_select, JP_AIDA_NK_dream_modelFit_Red_ref_SexMale_select, JP_AIDA_T_dream_modelFit_Red_ref_SexMale_select, JP_AIDA_B_dream_modelFit_Red_ref_SexMale_select) %>% reduce(full_join, by='Gene') %>% filter(Gene %in% TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_DEGs$Gene)

rownames(JP_AIDA_Allcelltypes_dream_modelFit_Red_ref_SexMale_df_representative_genes) <- JP_AIDA_Allcelltypes_dream_modelFit_Red_ref_SexMale_df_representative_genes$Gene
JP_AIDA_Allcelltypes_dream_modelFit_Red_ref_SexMale_df_representative_genes_mtx <- JP_AIDA_Allcelltypes_dream_modelFit_Red_ref_SexMale_df_representative_genes %>% select(-Gene) %>% replace(is.na(.), 0) %>% as.matrix()
```

#6. CD16+ NK: Sex differences among populations
```{r}
#NK
NK_ILC_Output_PATH <- "/mnt/icbs_shared_storage_general_2022/Benjamaporn/Analysis2025/Analysis202511/NK_ILC_integration"
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_NK_ILC_QC_RPCA_joined_clustered <- readRDS(file.path(NK_ILC_Output_PATH, "/AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_NK_ILC_QC_RPCA_joined_clustered.rds"))
```

```{r}
AIDA_ethnicity <- c("Thai", "Malay", "Chinese", "Indian", "Korean", "Japanese")
NK_ILC_merged_RPCA_joined_clustered_Ethnicity_CD16NK <- list()
AIDA_NK_ILC_metadata_perDCP_CD16NK <- list()
NK_ILC_merged_RPCA_joined_clustered_Ethnicity_CD16NK_count <- list()
NK_ILC_merged_RPCA_joined_clustered_Ethnicity_CD16NK_countdata <- list()
NK_ILC_merged_RPCA_joined_clustered_Ethnicity_CD16NK_DGE <- list()
NK_ILC_merged_RPCA_joined_clustered_Ethnicity_CD16NK_DGE_TMM <- list()

for (i in AIDA_ethnicity) {
#Metadata
NK_ILC_merged_RPCA_joined_clustered_Ethnicity_CD16NK[[i]] <- subset(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_NK_ILC_QC_RPCA_joined_clustered, subset = Ethnicity == i & Author_Annotation == "CD16+_NK")
AIDA_NK_ILC_metadata_perDCP_CD16NK[[i]] <- select(NK_ILC_merged_RPCA_joined_clustered_Ethnicity_CD16NK[[i]]@meta.data, Genotyping_ID, Ethnicity, DCP_ID, Age, Sex, Author_Annotation, Blue_ref, Red_ref, Orange_ref) %>%
distinct(DCP_ID, .keep_all = TRUE)  
row.names(AIDA_NK_ILC_metadata_perDCP_CD16NK[[i]]) <- NULL
AIDA_NK_ILC_metadata_perDCP_CD16NK[[i]]$DCP_ID <- gsub("_", "-", AIDA_NK_ILC_metadata_perDCP_CD16NK[[i]]$DCP_ID)
NK_ILC_merged_RPCA_joined_clustered_Ethnicity_CD16NK_count[[i]] <- AggregateExpression(NK_ILC_merged_RPCA_joined_clustered_Ethnicity_CD16NK[[i]], group.by = "DCP_ID", return.seurat = F)
NK_ILC_merged_RPCA_joined_clustered_Ethnicity_CD16NK_countdata[[i]] <- as.matrix(NK_ILC_merged_RPCA_joined_clustered_Ethnicity_CD16NK_count[[i]][["RNA"]])
NK_ILC_merged_RPCA_joined_clustered_Ethnicity_CD16NK_DGE[[i]] <- DGEList(NK_ILC_merged_RPCA_joined_clustered_Ethnicity_CD16NK_countdata[[i]])
NK_ILC_merged_RPCA_joined_clustered_Ethnicity_CD16NK_DGE_TMM[[i]] <- calcNormFactors(NK_ILC_merged_RPCA_joined_clustered_Ethnicity_CD16NK_DGE[[i]])
}
```

```{r}
AIDA_Blue_ethnicity <- c("Thai", "Malay", "Chinese")
CD16NK_DCP_Blue_design <- list()
bulk_CD16NK_Blue_DGE_TMM_vobjGenes <- list()
bulk_CD16NK_Blue_DGE_TMM_vobjGenes_mtx <- list()
bulk_CD16NK_Blue_DGE_TMM_vobjGenes_mtx_df <- list()
AIDA_NK_ILC_metadata_perDCP_Blue_CD16NK_merge <- list()
for (i in AIDA_Blue_ethnicity){
CD16NK_DCP_Blue_design[[i]] <- model.matrix(~ Sex + Blue_ref + Age, AIDA_NK_ILC_metadata_perDCP_CD16NK[[i]])
bulk_CD16NK_Blue_DGE_TMM_vobjGenes[[i]] <- voom(NK_ILC_merged_RPCA_joined_clustered_Ethnicity_CD16NK_DGE_TMM[[i]], CD16NK_DCP_Blue_design[[i]])
bulk_CD16NK_Blue_DGE_TMM_vobjGenes_mtx[[i]] <- as.matrix(bulk_CD16NK_Blue_DGE_TMM_vobjGenes[[i]])
bulk_CD16NK_Blue_DGE_TMM_vobjGenes_mtx_df[[i]] <- data.frame(t(bulk_CD16NK_Blue_DGE_TMM_vobjGenes_mtx[[i]]))
bulk_CD16NK_Blue_DGE_TMM_vobjGenes_mtx_df[[i]]$DCP_ID <- rownames(bulk_CD16NK_Blue_DGE_TMM_vobjGenes_mtx_df[[i]])
AIDA_NK_ILC_metadata_perDCP_Blue_CD16NK_merge[[i]] <- merge(bulk_CD16NK_Blue_DGE_TMM_vobjGenes_mtx_df[[i]], AIDA_NK_ILC_metadata_perDCP_CD16NK[[i]], by = "DCP_ID")  %>% 
  mutate(Gender = case_when(Sex == "Male" ~ "XY",
                          Sex == "Female" ~ "XX"
                                   )) %>% 
  mutate(Country_eth = case_when(
                                 Ethnicity == "Malay" ~ "SG_Malay",
                                 Ethnicity == "Thai" ~ "Thai",
                                 Ethnicity == "Chinese" ~ "SG_Chinese"
                                   ))
}

AIDA_Red_ethnicity <- c("Korean", "Japanese")
CD16NK_DCP_Red_design <- list()
bulk_CD16NK_Red_DGE_TMM_vobjGenes <- list()
bulk_CD16NK_Red_DGE_TMM_vobjGenes_mtx <- list()
bulk_CD16NK_Red_DGE_TMM_vobjGenes_mtx_df <- list()
AIDA_NK_ILC_metadata_perDCP_Red_CD16NK_merge <- list()
for (i in AIDA_Red_ethnicity){
CD16NK_DCP_Red_design[[i]] <- model.matrix(~ Sex + Red_ref + Age, AIDA_NK_ILC_metadata_perDCP_CD16NK[[i]])
bulk_CD16NK_Red_DGE_TMM_vobjGenes[[i]] <- voom(NK_ILC_merged_RPCA_joined_clustered_Ethnicity_CD16NK_DGE_TMM[[i]], CD16NK_DCP_Red_design[[i]])
bulk_CD16NK_Red_DGE_TMM_vobjGenes_mtx[[i]] <- as.matrix(bulk_CD16NK_Red_DGE_TMM_vobjGenes[[i]])
bulk_CD16NK_Red_DGE_TMM_vobjGenes_mtx_df[[i]] <- data.frame(t(bulk_CD16NK_Red_DGE_TMM_vobjGenes_mtx[[i]]))
bulk_CD16NK_Red_DGE_TMM_vobjGenes_mtx_df[[i]]$DCP_ID <- rownames(bulk_CD16NK_Red_DGE_TMM_vobjGenes_mtx_df[[i]])
AIDA_NK_ILC_metadata_perDCP_Red_CD16NK_merge[[i]] <- merge(bulk_CD16NK_Red_DGE_TMM_vobjGenes_mtx_df[[i]], AIDA_NK_ILC_metadata_perDCP_CD16NK[[i]], by = "DCP_ID")  %>% 
  mutate(Gender = case_when(Sex == "Male" ~ "XY",
                          Sex == "Female" ~ "XX"
                                   )) %>% 
  mutate(Country_eth = case_when(
                                 Ethnicity == "Korean" ~ "Korean",
                                 Ethnicity == "Japanese" ~ "Japanese"
                                   ))
}

#Orange
CD16NK_DCP_Orange_design <- model.matrix(~ Sex + Orange_ref + Age, AIDA_NK_ILC_metadata_perDCP_CD16NK[["Indian"]])
bulk_CD16NK_Orange_DGE_TMM_vobjGenes <- voom(NK_ILC_merged_RPCA_joined_clustered_Ethnicity_CD16NK_DGE_TMM[["Indian"]], CD16NK_DCP_Orange_design)
bulk_CD16NK_Orange_DGE_TMM_vobjGenes_mtx <- as.matrix(bulk_CD16NK_Orange_DGE_TMM_vobjGenes)
bulk_CD16NK_Orange_DGE_TMM_vobjGenes_mtx_df <- data.frame(t(bulk_CD16NK_Orange_DGE_TMM_vobjGenes_mtx))
bulk_CD16NK_Orange_DGE_TMM_vobjGenes_mtx_df$DCP_ID <- rownames(bulk_CD16NK_Orange_DGE_TMM_vobjGenes_mtx_df)
AIDA_NK_ILC_metadata_perDCP_Orange_CD16NK_merge <- merge(bulk_CD16NK_Orange_DGE_TMM_vobjGenes_mtx_df, AIDA_NK_ILC_metadata_perDCP_CD16NK[["Indian"]], by = "DCP_ID")  %>% 
  mutate(Gender = case_when(Sex == "Male" ~ "XY",
                          Sex == "Female" ~ "XX"
                                   )) %>% 
  mutate(Country_eth = case_when(Ethnicity == "Indian" ~ "SG_Indian"
                                   ))
```

```{r}
#Select only genes for comparison
bulk_TH_AIDA_NK_ILC_metadata_perDCP_CD16NK_merge <- select(AIDA_NK_ILC_metadata_perDCP_Blue_CD16NK_merge[["Thai"]], DCP_ID, Sex, Gender, Country_eth, XIST, RPS4Y1, DAPK1) %>%
mutate(Ethnicity = case_when(startsWith(DCP_ID, "TH") ~ "Thai"))
bulk_SG_Malay_AIDA_NK_ILC_metadata_perDCP_CD16NK_merge <- select(AIDA_NK_ILC_metadata_perDCP_Blue_CD16NK_merge[["Malay"]], DCP_ID, Sex, Gender, Country_eth, XIST, RPS4Y1, DAPK1) %>%
mutate(Ethnicity = case_when(startsWith(DCP_ID, "SG") ~ "SG_Malay"))
bulk_SG_Chinese_AIDA_NK_ILC_metadata_perDCP_CD16NK_merge <- select(AIDA_NK_ILC_metadata_perDCP_Blue_CD16NK_merge[["Chinese"]], DCP_ID,Sex, Gender, Country_eth, XIST, RPS4Y1, DAPK1) %>%
mutate(Ethnicity = case_when(startsWith(DCP_ID, "SG") ~ "SG_Chinese"))
bulk_SG_Indian_AIDA_NK_ILC_metadata_perDCP_CD16NK_merge <- select(AIDA_NK_ILC_metadata_perDCP_Orange_CD16NK_merge, DCP_ID, Sex, Gender, Country_eth, XIST, RPS4Y1, DAPK1) %>%
mutate(Ethnicity = case_when(startsWith(DCP_ID, "SG") ~ "SG_Indian"))
bulk_Korean_AIDA_NK_ILC_metadata_perDCP_CD16NK_merge <- select(AIDA_NK_ILC_metadata_perDCP_Red_CD16NK_merge[["Korean"]], DCP_ID, Sex, Gender, Country_eth, XIST, RPS4Y1, DAPK1) %>%
mutate(Ethnicity = case_when(startsWith(DCP_ID, "KR") ~ "Korean"))
bulk_Japanese_AIDA_NK_ILC_metadata_perDCP_CD16NK_merge <- select(AIDA_NK_ILC_metadata_perDCP_Red_CD16NK_merge[["Japanese"]], DCP_ID, Sex, Gender, Country_eth, XIST, RPS4Y1, DAPK1) %>%
mutate(Ethnicity = case_when(startsWith(DCP_ID, "JP") ~ "Japanese"))

#Merge df
All_AIDA_population_CD16NK_merge_genes <- rbind(bulk_TH_AIDA_NK_ILC_metadata_perDCP_CD16NK_merge, bulk_SG_Malay_AIDA_NK_ILC_metadata_perDCP_CD16NK_merge, bulk_SG_Chinese_AIDA_NK_ILC_metadata_perDCP_CD16NK_merge,
bulk_SG_Indian_AIDA_NK_ILC_metadata_perDCP_CD16NK_merge,
bulk_Korean_AIDA_NK_ILC_metadata_perDCP_CD16NK_merge,
bulk_Japanese_AIDA_NK_ILC_metadata_perDCP_CD16NK_merge)
```

```{r fig.height=4.5, fig.width=12}
Gender_col_dict <- c("XY" = "#298ccc", "XX" ="#c90076")
#Plotting
All_AIDA_population_CD16NK_merge_genes$Gender <- factor(All_AIDA_population_CD16NK_merge_genes$Gender, levels = c("XX", "XY"))
All_AIDA_population_CD16NK_merge_genes$Country_eth <- factor(All_AIDA_population_CD16NK_merge_genes$Country_eth, levels = c("SG_Indian", "SG_Malay", "Thai", "SG_Chinese", "Korean", "Japanese"))
#Add p-value manual
bulk_ALL_AIDA_CD16NK_stat_DAPK1_sexLMM <- data.frame(
    Country_eth= c("SG_Indian", "SG_Malay", "Thai", "SG_Chinese", "Korean", "Japanese"),
    group1 = c("XX", "XX", "XX", "XX", "XX", "XX"),
    group2 = c("XY", "XY", "XY", "XY", "XY", "XY"),
    p.adj = c("ns", "ns", "**", "ns", "ns", "ns")
)
bulk_ALL_AIDA_CD16NK_stat_DAPK1_sexLMM$Country_eth <- factor(bulk_ALL_AIDA_CD16NK_stat_DAPK1_sexLMM$Country_eth, levels = c("SG_Indian", "SG_Malay", "Thai", "SG_Chinese", "Korean", "Japanese"))

#Fig. 4E
bulk_ALL_AIDA_CD16NK_integrated_passQC_DCP_hvg_DGE_vobjGenes_mtx_df_merge_DAPK1_Sex_plot <- ggboxplot(All_AIDA_population_CD16NK_merge_genes, x="Gender", y="DAPK1", outlier.shape=NA, size=0.5, facet.by = "Country_eth") +
facet_grid(.~Country_eth) +
geom_jitter(width = 0.08, size= 0.8, height=0, aes(color=Gender))  +
ylab("logcpm" ~ italic("DAPK1")) +
xlab("")+
scale_color_manual(values=Gender_col_dict)+
theme(text = element_text(size = 16), axis.text.x = element_blank(), legend.position="bottom", strip.background = element_rect(colour="black", fill="white", size=1.5, linetype="solid")) + 
guides(color = guide_legend(override.aes = list(size = 8)))  + 
stat_pvalue_manual(bulk_ALL_AIDA_CD16NK_stat_DAPK1_sexLMM, label = "p.adj", hide.ns = T, y.position = 6, size = 4) +
labs(title = "CD16+ NK cells") 
bulk_ALL_AIDA_CD16NK_integrated_passQC_DCP_hvg_DGE_vobjGenes_mtx_df_merge_DAPK1_Sex_plot
```

#7. Pathway analysis for sex differences
```{r}
set.seed(123456)
library(tidyverse) 
library(RColorBrewer) 
library(fgsea)
library(msigdbr)
```

```{r}
#Retrieve the gene sets data frame 
human_gene_sets_df <- msigdbr(db_species = "HS", species = 'Homo sapiens', category = 'H')
human_gene_sets_df_C1 <- msigdbr(db_species = "HS", species = 'Homo sapiens', category = 'C1')

#Convert to named list format required by fgsea
human_gene_sets_gsea <- human_gene_sets_df %>% split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(human_gene_sets_gsea), "gene sets\n"))

#Pathways
human_gene_sets_gsea  <- human_gene_sets_gsea[!is.na(names(human_gene_sets_gsea))]
```

```{r}
#Myeloid_pDC
TH_dge_blue_ref_SexMale_Myeloid_pDC <- list()
TH_dge_blue_ref_SexMale_Myeloid_pDC_rankings <- list()
TH_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres <- list()
TH_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
TH_dge_blue_ref_SexMale_Myeloid_pDC[[i]] <- data.frame(TH_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale[[i]])
TH_dge_blue_ref_SexMale_Myeloid_pDC_rankings[[i]] <- sign(TH_dge_blue_ref_SexMale_Myeloid_pDC[[i]]$logFC)*(-log10(TH_dge_blue_ref_SexMale_Myeloid_pDC[[i]]$adj.P.Val))
names(TH_dge_blue_ref_SexMale_Myeloid_pDC_rankings[[i]]) <- TH_dge_blue_ref_SexMale_Myeloid_pDC[[i]]$Gene
TH_dge_blue_ref_SexMale_Myeloid_pDC_rankings[[i]] <- sort(TH_dge_blue_ref_SexMale_Myeloid_pDC_rankings[[i]], decreasing = TRUE)
TH_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = TH_dge_blue_ref_SexMale_Myeloid_pDC_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
TH_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select[[i]] <- select(TH_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
TH_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select[[i]]$Author_Annotation <- i
}

#NK
NK_Author_Annotation_celltype <- c("NK_unknown", "CD16+_NK", "CD56+_NK")
TH_dge_blue_ref_SexMale_NK <- list()
TH_dge_blue_ref_SexMale_NK_rankings <- list()
TH_dge_blue_ref_SexMale_NK_GSEAres <- list()
TH_dge_blue_ref_SexMale_NK_GSEAres_select <- list()
for (i in NK_Author_Annotation_celltype) {
TH_dge_blue_ref_SexMale_NK[[i]] <- data.frame(TH_AIDA_NK_dream_modelFit_Blue_ref_SexMale[[i]])
TH_dge_blue_ref_SexMale_NK_rankings[[i]] <- sign(TH_dge_blue_ref_SexMale_NK[[i]]$logFC)*(-log10(TH_dge_blue_ref_SexMale_NK[[i]]$adj.P.Val))
names(TH_dge_blue_ref_SexMale_NK_rankings[[i]]) <- TH_dge_blue_ref_SexMale_NK[[i]]$Gene
TH_dge_blue_ref_SexMale_NK_rankings[[i]] <- sort(TH_dge_blue_ref_SexMale_NK_rankings[[i]], decreasing = TRUE)
TH_dge_blue_ref_SexMale_NK_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = TH_dge_blue_ref_SexMale_NK_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
TH_dge_blue_ref_SexMale_NK_GSEAres_select[[i]] <- select(TH_dge_blue_ref_SexMale_NK_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
TH_dge_blue_ref_SexMale_NK_GSEAres_select[[i]]$Author_Annotation <- i
}

#T
T_Author_Annotation_celltype <- c("T_unknown", "CD4+_T_unknown", "CD4+_T_naive", "CD4+_T_cm",  "CD4+_T_cyt", "CD4+_T_em", "CD8+_T_unknown", "CD8+_T_naive", "CD8+_T_GZMBhi", "CD8+_T_GZMKhi", "MAIT",   "gdT_GZMBhi", "gdT_GZMKhi",  "Treg")
TH_dge_blue_ref_SexMale_T <- list()
TH_dge_blue_ref_SexMale_T_rankings <- list()
TH_dge_blue_ref_SexMale_T_GSEAres <- list()
TH_dge_blue_ref_SexMale_T_GSEAres_select <- list()
for (i in T_Author_Annotation_celltype) {
TH_dge_blue_ref_SexMale_T[[i]] <- data.frame(TH_AIDA_T_dream_modelFit_Blue_ref_SexMale[[i]])
TH_dge_blue_ref_SexMale_T_rankings[[i]] <- sign(TH_dge_blue_ref_SexMale_T[[i]]$logFC)*(-log10(TH_dge_blue_ref_SexMale_T[[i]]$adj.P.Val))
names(TH_dge_blue_ref_SexMale_T_rankings[[i]]) <- TH_dge_blue_ref_SexMale_T[[i]]$Gene
TH_dge_blue_ref_SexMale_T_rankings[[i]] <- sort(TH_dge_blue_ref_SexMale_T_rankings[[i]], decreasing = TRUE)
TH_dge_blue_ref_SexMale_T_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = TH_dge_blue_ref_SexMale_T_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
TH_dge_blue_ref_SexMale_T_GSEAres_select[[i]] <- select(TH_dge_blue_ref_SexMale_T_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
TH_dge_blue_ref_SexMale_T_GSEAres_select[[i]]$Author_Annotation <- i
}

#B
B_Author_Annotation_celltype <- c("B_unknown", "naive_B", "memory_B_IGHMlo", "memory_B_IGHMhi")
TH_dge_blue_ref_SexMale_B <- list()
TH_dge_blue_ref_SexMale_B_rankings <- list()
TH_dge_blue_ref_SexMale_B_GSEAres <- list()
TH_dge_blue_ref_SexMale_B_GSEAres_select <- list()
for (i in B_Author_Annotation_celltype) {
TH_dge_blue_ref_SexMale_B[[i]] <- data.frame(TH_AIDA_B_dream_modelFit_Blue_ref_SexMale[[i]])
TH_dge_blue_ref_SexMale_B_rankings[[i]] <- sign(TH_dge_blue_ref_SexMale_B[[i]]$logFC)*(-log10(TH_dge_blue_ref_SexMale_B[[i]]$adj.P.Val))
names(TH_dge_blue_ref_SexMale_B_rankings[[i]]) <- TH_dge_blue_ref_SexMale_B[[i]]$Gene
TH_dge_blue_ref_SexMale_B_rankings[[i]] <- sort(TH_dge_blue_ref_SexMale_B_rankings[[i]], decreasing = TRUE)
TH_dge_blue_ref_SexMale_B_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = TH_dge_blue_ref_SexMale_B_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
TH_dge_blue_ref_SexMale_B_GSEAres_select[[i]] <- select(TH_dge_blue_ref_SexMale_B_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
TH_dge_blue_ref_SexMale_B_GSEAres_select[[i]]$Author_Annotation <- i
}
```

```{r}
#Merge all cell types to get all GSEA results
TH_dge_blue_ref_SexMale_All_celltypes_GSEAres_df <- rbind(TH_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select[["Myeloid_unknown"]], TH_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select[["CD14+_Monocyte"]], TH_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select[["CD16+_Monocyte"]], TH_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select[["cDC2"]], TH_dge_blue_ref_SexMale_NK_GSEAres_select[["NK_unknown"]],
TH_dge_blue_ref_SexMale_NK_GSEAres_select[["CD16+_NK"]], TH_dge_blue_ref_SexMale_NK_GSEAres_select[["CD56+_NK"]],
TH_dge_blue_ref_SexMale_T_GSEAres_select[["T_unknown"]], TH_dge_blue_ref_SexMale_T_GSEAres_select[["CD4+_T_unknown"]], TH_dge_blue_ref_SexMale_T_GSEAres_select[["CD4+_T_naive"]], TH_dge_blue_ref_SexMale_T_GSEAres_select[["CD4+_T_cm"]], TH_dge_blue_ref_SexMale_T_GSEAres_select[["CD4+_T_cyt"]], TH_dge_blue_ref_SexMale_T_GSEAres_select[["CD4+_T_em"]], TH_dge_blue_ref_SexMale_T_GSEAres_select[["CD8+_T_unknown"]], TH_dge_blue_ref_SexMale_T_GSEAres_select[["CD8+_T_naive"]], TH_dge_blue_ref_SexMale_T_GSEAres_select[["CD8+_T_GZMBhi"]], TH_dge_blue_ref_SexMale_T_GSEAres_select[["CD8+_T_GZMKhi"]], TH_dge_blue_ref_SexMale_T_GSEAres_select[["MAIT"]], TH_dge_blue_ref_SexMale_T_GSEAres_select[["gdT_GZMBhi"]], TH_dge_blue_ref_SexMale_T_GSEAres_select[["gdT_GZMKhi"]], TH_dge_blue_ref_SexMale_T_GSEAres_select[["Treg"]],
TH_dge_blue_ref_SexMale_B_GSEAres_select[["B_unknown"]], TH_dge_blue_ref_SexMale_B_GSEAres_select[["naive_B"]], TH_dge_blue_ref_SexMale_B_GSEAres_select[["memory_B_IGHMlo"]], TH_dge_blue_ref_SexMale_B_GSEAres_select[["memory_B_IGHMhi"]])
```

```{r}
Author_Annotation_celltype = c("Myeloid_unknown", "CD14+_Monocyte", "CD16+_Monocyte", "cDC2", "Platelet", "NK_unknown", "CD16+_NK", "CD56+_NK", "T_unknown", "CD4+_T_unknown", "CD4+_T_naive", "CD4+_T_cm",  "CD4+_T_cyt", "CD4+_T_em", "CD8+_T_unknown", "CD8+_T_naive", "CD8+_T_GZMBhi", "CD8+_T_GZMKhi", "MAIT",   "gdT_GZMBhi", "gdT_GZMKhi",  "Treg", "B_unknown", "naive_B", "memory_B_IGHMlo", "memory_B_IGHMhi")
```

```{r}
#Myeloid_pDC
SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC <- list()
SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC_rankings <- list()
SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres <- list()
SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC[[i]] <- data.frame(SG_Malay_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale[[i]])
SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC_rankings[[i]] <- sign(SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC[[i]]$logFC)*(-log10(SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC[[i]]$adj.P.Val))
names(SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC_rankings[[i]]) <- SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC[[i]]$Gene
SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC_rankings[[i]] <- sort(SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC_rankings[[i]], decreasing = TRUE)
SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select[[i]] <- select(SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select[[i]]$Author_Annotation <- i
}

#NK
SG_Malay_dge_blue_ref_SexMale_NK <- list()
SG_Malay_dge_blue_ref_SexMale_NK_rankings <- list()
SG_Malay_dge_blue_ref_SexMale_NK_GSEAres <- list()
SG_Malay_dge_blue_ref_SexMale_NK_GSEAres_select <- list()
for (i in NK_Author_Annotation_celltype) {
SG_Malay_dge_blue_ref_SexMale_NK[[i]] <- data.frame(SG_Malay_AIDA_NK_dream_modelFit_Blue_ref_SexMale[[i]])
SG_Malay_dge_blue_ref_SexMale_NK_rankings[[i]] <- sign(SG_Malay_dge_blue_ref_SexMale_NK[[i]]$logFC)*(-log10(SG_Malay_dge_blue_ref_SexMale_NK[[i]]$adj.P.Val))
names(SG_Malay_dge_blue_ref_SexMale_NK_rankings[[i]]) <- SG_Malay_dge_blue_ref_SexMale_NK[[i]]$Gene
SG_Malay_dge_blue_ref_SexMale_NK_rankings[[i]] <- sort(SG_Malay_dge_blue_ref_SexMale_NK_rankings[[i]], decreasing = TRUE)
SG_Malay_dge_blue_ref_SexMale_NK_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = SG_Malay_dge_blue_ref_SexMale_NK_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
SG_Malay_dge_blue_ref_SexMale_NK_GSEAres_select[[i]] <- select(SG_Malay_dge_blue_ref_SexMale_NK_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
SG_Malay_dge_blue_ref_SexMale_NK_GSEAres_select[[i]]$Author_Annotation <- i
}

#T
SG_Malay_dge_blue_ref_SexMale_T <- list()
SG_Malay_dge_blue_ref_SexMale_T_rankings <- list()
SG_Malay_dge_blue_ref_SexMale_T_GSEAres <- list()
SG_Malay_dge_blue_ref_SexMale_T_GSEAres_select <- list()
for (i in T_Author_Annotation_celltype) {
SG_Malay_dge_blue_ref_SexMale_T[[i]] <- data.frame(SG_Malay_AIDA_T_dream_modelFit_Blue_ref_SexMale[[i]])
SG_Malay_dge_blue_ref_SexMale_T_rankings[[i]] <- sign(SG_Malay_dge_blue_ref_SexMale_T[[i]]$logFC)*(-log10(SG_Malay_dge_blue_ref_SexMale_T[[i]]$adj.P.Val))
names(SG_Malay_dge_blue_ref_SexMale_T_rankings[[i]]) <- SG_Malay_dge_blue_ref_SexMale_T[[i]]$Gene
SG_Malay_dge_blue_ref_SexMale_T_rankings[[i]] <- sort(SG_Malay_dge_blue_ref_SexMale_T_rankings[[i]], decreasing = TRUE)
SG_Malay_dge_blue_ref_SexMale_T_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = SG_Malay_dge_blue_ref_SexMale_T_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
SG_Malay_dge_blue_ref_SexMale_T_GSEAres_select[[i]] <- select(SG_Malay_dge_blue_ref_SexMale_T_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
SG_Malay_dge_blue_ref_SexMale_T_GSEAres_select[[i]]$Author_Annotation <- i
}

#B
SG_Malay_dge_blue_ref_SexMale_B <- list()
SG_Malay_dge_blue_ref_SexMale_B_rankings <- list()
SG_Malay_dge_blue_ref_SexMale_B_GSEAres <- list()
SG_Malay_dge_blue_ref_SexMale_B_GSEAres_select <- list()
for (i in B_Author_Annotation_celltype) {
SG_Malay_dge_blue_ref_SexMale_B[[i]] <- data.frame(SG_Malay_AIDA_B_dream_modelFit_Blue_ref_SexMale[[i]])
SG_Malay_dge_blue_ref_SexMale_B_rankings[[i]] <- sign(SG_Malay_dge_blue_ref_SexMale_B[[i]]$logFC)*(-log10(SG_Malay_dge_blue_ref_SexMale_B[[i]]$adj.P.Val))
names(SG_Malay_dge_blue_ref_SexMale_B_rankings[[i]]) <- SG_Malay_dge_blue_ref_SexMale_B[[i]]$Gene
SG_Malay_dge_blue_ref_SexMale_B_rankings[[i]] <- sort(SG_Malay_dge_blue_ref_SexMale_B_rankings[[i]], decreasing = TRUE)
SG_Malay_dge_blue_ref_SexMale_B_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = SG_Malay_dge_blue_ref_SexMale_B_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
SG_Malay_dge_blue_ref_SexMale_B_GSEAres_select[[i]] <- select(SG_Malay_dge_blue_ref_SexMale_B_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
SG_Malay_dge_blue_ref_SexMale_B_GSEAres_select[[i]]$Author_Annotation <- i
}
```

```{r}
#Merge all cell types to get all GSEA results
SG_Malay_dge_blue_ref_SexMale_All_celltypes_GSEAres_df <- rbind(SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select[["Myeloid_unknown"]], SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select[["CD14+_Monocyte"]], SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select[["CD16+_Monocyte"]], SG_Malay_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select[["cDC2"]], SG_Malay_dge_blue_ref_SexMale_NK_GSEAres_select[["NK_unknown"]],
SG_Malay_dge_blue_ref_SexMale_NK_GSEAres_select[["CD16+_NK"]], SG_Malay_dge_blue_ref_SexMale_NK_GSEAres_select[["CD56+_NK"]],
SG_Malay_dge_blue_ref_SexMale_T_GSEAres_select[["T_unknown"]], SG_Malay_dge_blue_ref_SexMale_T_GSEAres_select[["CD4+_T_unknown"]], SG_Malay_dge_blue_ref_SexMale_T_GSEAres_select[["CD4+_T_naive"]], SG_Malay_dge_blue_ref_SexMale_T_GSEAres_select[["CD4+_T_cm"]], SG_Malay_dge_blue_ref_SexMale_T_GSEAres_select[["CD4+_T_cyt"]], SG_Malay_dge_blue_ref_SexMale_T_GSEAres_select[["CD4+_T_em"]], SG_Malay_dge_blue_ref_SexMale_T_GSEAres_select[["CD8+_T_unknown"]], SG_Malay_dge_blue_ref_SexMale_T_GSEAres_select[["CD8+_T_naive"]], SG_Malay_dge_blue_ref_SexMale_T_GSEAres_select[["CD8+_T_GZMBhi"]], SG_Malay_dge_blue_ref_SexMale_T_GSEAres_select[["CD8+_T_GZMKhi"]], SG_Malay_dge_blue_ref_SexMale_T_GSEAres_select[["MAIT"]], SG_Malay_dge_blue_ref_SexMale_T_GSEAres_select[["gdT_GZMBhi"]], SG_Malay_dge_blue_ref_SexMale_T_GSEAres_select[["gdT_GZMKhi"]], SG_Malay_dge_blue_ref_SexMale_T_GSEAres_select[["Treg"]],SG_Malay_dge_blue_ref_SexMale_B_GSEAres_select[["B_unknown"]], SG_Malay_dge_blue_ref_SexMale_B_GSEAres_select[["naive_B"]], SG_Malay_dge_blue_ref_SexMale_B_GSEAres_select[["memory_B_IGHMlo"]], SG_Malay_dge_blue_ref_SexMale_B_GSEAres_select[["memory_B_IGHMhi"]])
```

```{r}
#Myeloid_pDC
SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC <- list()
SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC_rankings <- list()
SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres <- list()
SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC[[i]] <- data.frame(SG_Chinese_AIDA_Myeloid_pDC_dream_modelFit_Blue_ref_SexMale[[i]])
SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC_rankings[[i]] <- sign(SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC[[i]]$logFC)*(-log10(SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC[[i]]$adj.P.Val))
names(SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC_rankings[[i]]) <- SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC[[i]]$Gene
SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC_rankings[[i]] <- sort(SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC_rankings[[i]], decreasing = TRUE)
SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select[[i]] <- select(SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select[[i]]$Author_Annotation <- i
}

#NK
SG_Chinese_dge_blue_ref_SexMale_NK <- list()
SG_Chinese_dge_blue_ref_SexMale_NK_rankings <- list()
SG_Chinese_dge_blue_ref_SexMale_NK_GSEAres <- list()
SG_Chinese_dge_blue_ref_SexMale_NK_GSEAres_select <- list()
for (i in NK_Author_Annotation_celltype) {
SG_Chinese_dge_blue_ref_SexMale_NK[[i]] <- data.frame(SG_Chinese_AIDA_NK_dream_modelFit_Blue_ref_SexMale[[i]])
SG_Chinese_dge_blue_ref_SexMale_NK_rankings[[i]] <- sign(SG_Chinese_dge_blue_ref_SexMale_NK[[i]]$logFC)*(-log10(SG_Chinese_dge_blue_ref_SexMale_NK[[i]]$adj.P.Val))
names(SG_Chinese_dge_blue_ref_SexMale_NK_rankings[[i]]) <- SG_Chinese_dge_blue_ref_SexMale_NK[[i]]$Gene
SG_Chinese_dge_blue_ref_SexMale_NK_rankings[[i]] <- sort(SG_Chinese_dge_blue_ref_SexMale_NK_rankings[[i]], decreasing = TRUE)
SG_Chinese_dge_blue_ref_SexMale_NK_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = SG_Chinese_dge_blue_ref_SexMale_NK_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
SG_Chinese_dge_blue_ref_SexMale_NK_GSEAres_select[[i]] <- select(SG_Chinese_dge_blue_ref_SexMale_NK_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
SG_Chinese_dge_blue_ref_SexMale_NK_GSEAres_select[[i]]$Author_Annotation <- i
}

#T
SG_Chinese_dge_blue_ref_SexMale_T <- list()
SG_Chinese_dge_blue_ref_SexMale_T_rankings <- list()
SG_Chinese_dge_blue_ref_SexMale_T_GSEAres <- list()
SG_Chinese_dge_blue_ref_SexMale_T_GSEAres_select <- list()
for (i in T_Author_Annotation_celltype) {
SG_Chinese_dge_blue_ref_SexMale_T[[i]] <- data.frame(SG_Chinese_AIDA_T_dream_modelFit_Blue_ref_SexMale[[i]])
SG_Chinese_dge_blue_ref_SexMale_T_rankings[[i]] <- sign(SG_Chinese_dge_blue_ref_SexMale_T[[i]]$logFC)*(-log10(SG_Chinese_dge_blue_ref_SexMale_T[[i]]$adj.P.Val))
names(SG_Chinese_dge_blue_ref_SexMale_T_rankings[[i]]) <- SG_Chinese_dge_blue_ref_SexMale_T[[i]]$Gene
SG_Chinese_dge_blue_ref_SexMale_T_rankings[[i]] <- sort(SG_Chinese_dge_blue_ref_SexMale_T_rankings[[i]], decreasing = TRUE)
SG_Chinese_dge_blue_ref_SexMale_T_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = SG_Chinese_dge_blue_ref_SexMale_T_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
SG_Chinese_dge_blue_ref_SexMale_T_GSEAres_select[[i]] <- select(SG_Chinese_dge_blue_ref_SexMale_T_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
SG_Chinese_dge_blue_ref_SexMale_T_GSEAres_select[[i]]$Author_Annotation <- i
}


#B
SG_Chinese_dge_blue_ref_SexMale_B <- list()
SG_Chinese_dge_blue_ref_SexMale_B_rankings <- list()
SG_Chinese_dge_blue_ref_SexMale_B_GSEAres <- list()
SG_Chinese_dge_blue_ref_SexMale_B_GSEAres_select <- list()
for (i in B_Author_Annotation_celltype) {
SG_Chinese_dge_blue_ref_SexMale_B[[i]] <- data.frame(SG_Chinese_AIDA_B_dream_modelFit_Blue_ref_SexMale[[i]])
SG_Chinese_dge_blue_ref_SexMale_B_rankings[[i]] <- sign(SG_Chinese_dge_blue_ref_SexMale_B[[i]]$logFC)*(-log10(SG_Chinese_dge_blue_ref_SexMale_B[[i]]$adj.P.Val))
names(SG_Chinese_dge_blue_ref_SexMale_B_rankings[[i]]) <- SG_Chinese_dge_blue_ref_SexMale_B[[i]]$Gene
SG_Chinese_dge_blue_ref_SexMale_B_rankings[[i]] <- sort(SG_Chinese_dge_blue_ref_SexMale_B_rankings[[i]], decreasing = TRUE)
SG_Chinese_dge_blue_ref_SexMale_B_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = SG_Chinese_dge_blue_ref_SexMale_B_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
SG_Chinese_dge_blue_ref_SexMale_B_GSEAres_select[[i]] <- select(SG_Chinese_dge_blue_ref_SexMale_B_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
SG_Chinese_dge_blue_ref_SexMale_B_GSEAres_select[[i]]$Author_Annotation <- i
}
```

```{r}
#Merge all cell types to get all GSEA results
SG_Chinese_dge_blue_ref_SexMale_All_celltypes_GSEAres_df <- rbind(SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select[["Myeloid_unknown"]], SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select[["CD14+_Monocyte"]], SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select[["CD16+_Monocyte"]], SG_Chinese_dge_blue_ref_SexMale_Myeloid_pDC_GSEAres_select[["cDC2"]], SG_Chinese_dge_blue_ref_SexMale_NK_GSEAres_select[["NK_unknown"]],
SG_Chinese_dge_blue_ref_SexMale_NK_GSEAres_select[["CD16+_NK"]], SG_Chinese_dge_blue_ref_SexMale_NK_GSEAres_select[["CD56+_NK"]],
SG_Chinese_dge_blue_ref_SexMale_T_GSEAres_select[["T_unknown"]], SG_Chinese_dge_blue_ref_SexMale_T_GSEAres_select[["CD4+_T_unknown"]], SG_Chinese_dge_blue_ref_SexMale_T_GSEAres_select[["CD4+_T_naive"]], SG_Chinese_dge_blue_ref_SexMale_T_GSEAres_select[["CD4+_T_cm"]], SG_Chinese_dge_blue_ref_SexMale_T_GSEAres_select[["CD4+_T_cyt"]], SG_Chinese_dge_blue_ref_SexMale_T_GSEAres_select[["CD4+_T_em"]], SG_Chinese_dge_blue_ref_SexMale_T_GSEAres_select[["CD8+_T_unknown"]], SG_Chinese_dge_blue_ref_SexMale_T_GSEAres_select[["CD8+_T_naive"]], SG_Chinese_dge_blue_ref_SexMale_T_GSEAres_select[["CD8+_T_GZMBhi"]], SG_Chinese_dge_blue_ref_SexMale_T_GSEAres_select[["CD8+_T_GZMKhi"]], SG_Chinese_dge_blue_ref_SexMale_T_GSEAres_select[["MAIT"]], SG_Chinese_dge_blue_ref_SexMale_T_GSEAres_select[["gdT_GZMBhi"]], SG_Chinese_dge_blue_ref_SexMale_T_GSEAres_select[["gdT_GZMKhi"]], SG_Chinese_dge_blue_ref_SexMale_T_GSEAres_select[["Treg"]], SG_Chinese_dge_blue_ref_SexMale_B_GSEAres_select[["B_unknown"]], SG_Chinese_dge_blue_ref_SexMale_B_GSEAres_select[["naive_B"]], SG_Chinese_dge_blue_ref_SexMale_B_GSEAres_select[["memory_B_IGHMlo"]], SG_Chinese_dge_blue_ref_SexMale_B_GSEAres_select[["memory_B_IGHMhi"]])
```

```{r}
#Myeloid_pDC
SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC <- list()
SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC_rankings <- list()
SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC_GSEAres <- list()
SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC_GSEAres_select <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC[[i]] <- data.frame(SG_Indian_AIDA_Myeloid_pDC_dream_modelFit_Orange_ref_SexMale[[i]])
SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC_rankings[[i]] <- sign(SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC[[i]]$logFC)*(-log10(SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC[[i]]$adj.P.Val))
names(SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC_rankings[[i]]) <- SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC[[i]]$Gene
SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC_rankings[[i]] <- sort(SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC_rankings[[i]], decreasing = TRUE)
SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC_GSEAres_select[[i]] <- select(SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC_GSEAres_select[[i]]$Author_Annotation <- i
}

#NK
SG_Indian_dge_Orange_ref_SexMale_NK <- list()
SG_Indian_dge_Orange_ref_SexMale_NK_rankings <- list()
SG_Indian_dge_Orange_ref_SexMale_NK_GSEAres <- list()
SG_Indian_dge_Orange_ref_SexMale_NK_GSEAres_select <- list()
for (i in NK_Author_Annotation_celltype) {
SG_Indian_dge_Orange_ref_SexMale_NK[[i]] <- data.frame(SG_Indian_AIDA_NK_dream_modelFit_Orange_ref_SexMale[[i]])
SG_Indian_dge_Orange_ref_SexMale_NK_rankings[[i]] <- sign(SG_Indian_dge_Orange_ref_SexMale_NK[[i]]$logFC)*(-log10(SG_Indian_dge_Orange_ref_SexMale_NK[[i]]$adj.P.Val))
names(SG_Indian_dge_Orange_ref_SexMale_NK_rankings[[i]]) <- SG_Indian_dge_Orange_ref_SexMale_NK[[i]]$Gene
SG_Indian_dge_Orange_ref_SexMale_NK_rankings[[i]] <- sort(SG_Indian_dge_Orange_ref_SexMale_NK_rankings[[i]], decreasing = TRUE)
SG_Indian_dge_Orange_ref_SexMale_NK_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = SG_Indian_dge_Orange_ref_SexMale_NK_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
SG_Indian_dge_Orange_ref_SexMale_NK_GSEAres_select[[i]] <- select(SG_Indian_dge_Orange_ref_SexMale_NK_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
SG_Indian_dge_Orange_ref_SexMale_NK_GSEAres_select[[i]]$Author_Annotation <- i
}

#T
SG_Indian_dge_Orange_ref_SexMale_T <- list()
SG_Indian_dge_Orange_ref_SexMale_T_rankings <- list()
SG_Indian_dge_Orange_ref_SexMale_T_GSEAres <- list()
SG_Indian_dge_Orange_ref_SexMale_T_GSEAres_select <- list()
for (i in T_Author_Annotation_celltype) {
SG_Indian_dge_Orange_ref_SexMale_T[[i]] <- data.frame(SG_Indian_AIDA_T_dream_modelFit_Orange_ref_SexMale[[i]])
SG_Indian_dge_Orange_ref_SexMale_T_rankings[[i]] <- sign(SG_Indian_dge_Orange_ref_SexMale_T[[i]]$logFC)*(-log10(SG_Indian_dge_Orange_ref_SexMale_T[[i]]$adj.P.Val))
names(SG_Indian_dge_Orange_ref_SexMale_T_rankings[[i]]) <- SG_Indian_dge_Orange_ref_SexMale_T[[i]]$Gene
SG_Indian_dge_Orange_ref_SexMale_T_rankings[[i]] <- sort(SG_Indian_dge_Orange_ref_SexMale_T_rankings[[i]], decreasing = TRUE)
SG_Indian_dge_Orange_ref_SexMale_T_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = SG_Indian_dge_Orange_ref_SexMale_T_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
SG_Indian_dge_Orange_ref_SexMale_T_GSEAres_select[[i]] <- select(SG_Indian_dge_Orange_ref_SexMale_T_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
SG_Indian_dge_Orange_ref_SexMale_T_GSEAres_select[[i]]$Author_Annotation <- i
}


#B
SG_Indian_dge_Orange_ref_SexMale_B <- list()
SG_Indian_dge_Orange_ref_SexMale_B_rankings <- list()
SG_Indian_dge_Orange_ref_SexMale_B_GSEAres <- list()
SG_Indian_dge_Orange_ref_SexMale_B_GSEAres_select <- list()
for (i in B_Author_Annotation_celltype) {
SG_Indian_dge_Orange_ref_SexMale_B[[i]] <- data.frame(SG_Indian_AIDA_B_dream_modelFit_Orange_ref_SexMale[[i]])
SG_Indian_dge_Orange_ref_SexMale_B_rankings[[i]] <- sign(SG_Indian_dge_Orange_ref_SexMale_B[[i]]$logFC)*(-log10(SG_Indian_dge_Orange_ref_SexMale_B[[i]]$adj.P.Val))
names(SG_Indian_dge_Orange_ref_SexMale_B_rankings[[i]]) <- SG_Indian_dge_Orange_ref_SexMale_B[[i]]$Gene
SG_Indian_dge_Orange_ref_SexMale_B_rankings[[i]] <- sort(SG_Indian_dge_Orange_ref_SexMale_B_rankings[[i]], decreasing = TRUE)
SG_Indian_dge_Orange_ref_SexMale_B_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = SG_Indian_dge_Orange_ref_SexMale_B_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
SG_Indian_dge_Orange_ref_SexMale_B_GSEAres_select[[i]] <- select(SG_Indian_dge_Orange_ref_SexMale_B_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
SG_Indian_dge_Orange_ref_SexMale_B_GSEAres_select[[i]]$Author_Annotation <- i
}
```

```{r}
#Merge all cell types to get all GSEA results
SG_Indian_dge_Orange_ref_SexMale_All_celltypes_GSEAres_df <- rbind(SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC_GSEAres_select[["Myeloid_unknown"]], SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC_GSEAres_select[["CD14+_Monocyte"]], SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC_GSEAres_select[["CD16+_Monocyte"]], SG_Indian_dge_Orange_ref_SexMale_Myeloid_pDC_GSEAres_select[["cDC2"]], SG_Indian_dge_Orange_ref_SexMale_NK_GSEAres_select[["NK_unknown"]],
SG_Indian_dge_Orange_ref_SexMale_NK_GSEAres_select[["CD16+_NK"]], SG_Indian_dge_Orange_ref_SexMale_NK_GSEAres_select[["CD56+_NK"]],
SG_Indian_dge_Orange_ref_SexMale_T_GSEAres_select[["T_unknown"]], SG_Indian_dge_Orange_ref_SexMale_T_GSEAres_select[["CD4+_T_unknown"]], SG_Indian_dge_Orange_ref_SexMale_T_GSEAres_select[["CD4+_T_naive"]], SG_Indian_dge_Orange_ref_SexMale_T_GSEAres_select[["CD4+_T_cm"]], SG_Indian_dge_Orange_ref_SexMale_T_GSEAres_select[["CD4+_T_cyt"]], SG_Indian_dge_Orange_ref_SexMale_T_GSEAres_select[["CD4+_T_em"]], SG_Indian_dge_Orange_ref_SexMale_T_GSEAres_select[["CD8+_T_unknown"]], SG_Indian_dge_Orange_ref_SexMale_T_GSEAres_select[["CD8+_T_naive"]], SG_Indian_dge_Orange_ref_SexMale_T_GSEAres_select[["CD8+_T_GZMBhi"]], SG_Indian_dge_Orange_ref_SexMale_T_GSEAres_select[["CD8+_T_GZMKhi"]], SG_Indian_dge_Orange_ref_SexMale_T_GSEAres_select[["MAIT"]], SG_Indian_dge_Orange_ref_SexMale_T_GSEAres_select[["gdT_GZMBhi"]], SG_Indian_dge_Orange_ref_SexMale_T_GSEAres_select[["gdT_GZMKhi"]], SG_Indian_dge_Orange_ref_SexMale_T_GSEAres_select[["Treg"]],
SG_Indian_dge_Orange_ref_SexMale_B_GSEAres_select[["B_unknown"]],
SG_Indian_dge_Orange_ref_SexMale_B_GSEAres_select[["naive_B"]], SG_Indian_dge_Orange_ref_SexMale_B_GSEAres_select[["memory_B_IGHMlo"]], SG_Indian_dge_Orange_ref_SexMale_B_GSEAres_select[["memory_B_IGHMhi"]])
```

```{r}
#Myeloid_pDC
Korean_dge_Red_ref_SexMale_Myeloid_pDC <- list()
Korean_dge_Red_ref_SexMale_Myeloid_pDC_rankings <- list()
Korean_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres <- list()
Korean_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres_select <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
Korean_dge_Red_ref_SexMale_Myeloid_pDC[[i]] <- data.frame(KR_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale[[i]])
Korean_dge_Red_ref_SexMale_Myeloid_pDC_rankings[[i]] <- sign(Korean_dge_Red_ref_SexMale_Myeloid_pDC[[i]]$logFC)*(-log10(Korean_dge_Red_ref_SexMale_Myeloid_pDC[[i]]$adj.P.Val))
names(Korean_dge_Red_ref_SexMale_Myeloid_pDC_rankings[[i]]) <- Korean_dge_Red_ref_SexMale_Myeloid_pDC[[i]]$Gene
Korean_dge_Red_ref_SexMale_Myeloid_pDC_rankings[[i]] <- sort(Korean_dge_Red_ref_SexMale_Myeloid_pDC_rankings[[i]], decreasing = TRUE)
Korean_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = Korean_dge_Red_ref_SexMale_Myeloid_pDC_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
Korean_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres_select[[i]] <- select(Korean_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
Korean_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres_select[[i]]$Author_Annotation <- i
}

#NK
Korean_dge_Red_ref_SexMale_NK <- list()
Korean_dge_Red_ref_SexMale_NK_rankings <- list()
Korean_dge_Red_ref_SexMale_NK_GSEAres <- list()
Korean_dge_Red_ref_SexMale_NK_GSEAres_select <- list()
for (i in NK_Author_Annotation_celltype) {
Korean_dge_Red_ref_SexMale_NK[[i]] <- data.frame(KR_AIDA_NK_dream_modelFit_Red_ref_SexMale[[i]])
Korean_dge_Red_ref_SexMale_NK_rankings[[i]] <- sign(Korean_dge_Red_ref_SexMale_NK[[i]]$logFC)*(-log10(Korean_dge_Red_ref_SexMale_NK[[i]]$adj.P.Val))
names(Korean_dge_Red_ref_SexMale_NK_rankings[[i]]) <- Korean_dge_Red_ref_SexMale_NK[[i]]$Gene
Korean_dge_Red_ref_SexMale_NK_rankings[[i]] <- sort(Korean_dge_Red_ref_SexMale_NK_rankings[[i]], decreasing = TRUE)
Korean_dge_Red_ref_SexMale_NK_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = Korean_dge_Red_ref_SexMale_NK_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
Korean_dge_Red_ref_SexMale_NK_GSEAres_select[[i]] <- select(Korean_dge_Red_ref_SexMale_NK_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
Korean_dge_Red_ref_SexMale_NK_GSEAres_select[[i]]$Author_Annotation <- i
}

#T
Korean_dge_Red_ref_SexMale_T <- list()
Korean_dge_Red_ref_SexMale_T_rankings <- list()
Korean_dge_Red_ref_SexMale_T_GSEAres <- list()
Korean_dge_Red_ref_SexMale_T_GSEAres_select <- list()
for (i in T_Author_Annotation_celltype) {
Korean_dge_Red_ref_SexMale_T[[i]] <- data.frame(KR_AIDA_T_dream_modelFit_Red_ref_SexMale[[i]])
Korean_dge_Red_ref_SexMale_T_rankings[[i]] <- sign(Korean_dge_Red_ref_SexMale_T[[i]]$logFC)*(-log10(Korean_dge_Red_ref_SexMale_T[[i]]$adj.P.Val))
names(Korean_dge_Red_ref_SexMale_T_rankings[[i]]) <- Korean_dge_Red_ref_SexMale_T[[i]]$Gene
Korean_dge_Red_ref_SexMale_T_rankings[[i]] <- sort(Korean_dge_Red_ref_SexMale_T_rankings[[i]], decreasing = TRUE)
Korean_dge_Red_ref_SexMale_T_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = Korean_dge_Red_ref_SexMale_T_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
Korean_dge_Red_ref_SexMale_T_GSEAres_select[[i]] <- select(Korean_dge_Red_ref_SexMale_T_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
Korean_dge_Red_ref_SexMale_T_GSEAres_select[[i]]$Author_Annotation <- i
}


#B
Korean_dge_Red_ref_SexMale_B <- list()
Korean_dge_Red_ref_SexMale_B_rankings <- list()
Korean_dge_Red_ref_SexMale_B_GSEAres <- list()
Korean_dge_Red_ref_SexMale_B_GSEAres_select <- list()
for (i in B_Author_Annotation_celltype) {
Korean_dge_Red_ref_SexMale_B[[i]] <- data.frame(KR_AIDA_B_dream_modelFit_Red_ref_SexMale[[i]])
Korean_dge_Red_ref_SexMale_B_rankings[[i]] <- sign(Korean_dge_Red_ref_SexMale_B[[i]]$logFC)*(-log10(Korean_dge_Red_ref_SexMale_B[[i]]$adj.P.Val))
names(Korean_dge_Red_ref_SexMale_B_rankings[[i]]) <- Korean_dge_Red_ref_SexMale_B[[i]]$Gene
Korean_dge_Red_ref_SexMale_B_rankings[[i]] <- sort(Korean_dge_Red_ref_SexMale_B_rankings[[i]], decreasing = TRUE)
Korean_dge_Red_ref_SexMale_B_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = Korean_dge_Red_ref_SexMale_B_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
Korean_dge_Red_ref_SexMale_B_GSEAres_select[[i]] <- select(Korean_dge_Red_ref_SexMale_B_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
Korean_dge_Red_ref_SexMale_B_GSEAres_select[[i]]$Author_Annotation <- i
}
```

```{r}
#Merge all cell types to get all GSEA results
Korean_dge_Red_ref_SexMale_All_celltypes_GSEAres_df <- rbind(Korean_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres_select[["Myeloid_unknown"]], Korean_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres_select[["CD14+_Monocyte"]], Korean_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres_select[["CD16+_Monocyte"]], Korean_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres_select[["cDC2"]], Korean_dge_Red_ref_SexMale_NK_GSEAres_select[["NK_unknown"]],
Korean_dge_Red_ref_SexMale_NK_GSEAres_select[["CD16+_NK"]], Korean_dge_Red_ref_SexMale_NK_GSEAres_select[["CD56+_NK"]],
Korean_dge_Red_ref_SexMale_T_GSEAres_select[["T_unknown"]], Korean_dge_Red_ref_SexMale_T_GSEAres_select[["CD4+_T_unknown"]], Korean_dge_Red_ref_SexMale_T_GSEAres_select[["CD4+_T_naive"]], Korean_dge_Red_ref_SexMale_T_GSEAres_select[["CD4+_T_cm"]], Korean_dge_Red_ref_SexMale_T_GSEAres_select[["CD4+_T_cyt"]], Korean_dge_Red_ref_SexMale_T_GSEAres_select[["CD4+_T_em"]], Korean_dge_Red_ref_SexMale_T_GSEAres_select[["CD8+_T_unknown"]], Korean_dge_Red_ref_SexMale_T_GSEAres_select[["CD8+_T_naive"]], Korean_dge_Red_ref_SexMale_T_GSEAres_select[["CD8+_T_GZMBhi"]], Korean_dge_Red_ref_SexMale_T_GSEAres_select[["CD8+_T_GZMKhi"]], Korean_dge_Red_ref_SexMale_T_GSEAres_select[["MAIT"]], Korean_dge_Red_ref_SexMale_T_GSEAres_select[["gdT_GZMBhi"]], Korean_dge_Red_ref_SexMale_T_GSEAres_select[["gdT_GZMKhi"]], Korean_dge_Red_ref_SexMale_T_GSEAres_select[["Treg"]],
Korean_dge_Red_ref_SexMale_B_GSEAres_select[["B_unknown"]], Korean_dge_Red_ref_SexMale_B_GSEAres_select[["naive_B"]], Korean_dge_Red_ref_SexMale_B_GSEAres_select[["memory_B_IGHMlo"]], Korean_dge_Red_ref_SexMale_B_GSEAres_select[["memory_B_IGHMhi"]])
```

```{r}
#Myeloid_pDC
Japanese_dge_Red_ref_SexMale_Myeloid_pDC <- list()
Japanese_dge_Red_ref_SexMale_Myeloid_pDC_rankings <- list()
Japanese_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres <- list()
Japanese_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres_select <- list()
for (i in Myeloid_pDC_Author_Annotation_celltype) {
Japanese_dge_Red_ref_SexMale_Myeloid_pDC[[i]] <- data.frame(JP_AIDA_Myeloid_pDC_dream_modelFit_Red_ref_SexMale[[i]])
Japanese_dge_Red_ref_SexMale_Myeloid_pDC_rankings[[i]] <- sign(Japanese_dge_Red_ref_SexMale_Myeloid_pDC[[i]]$logFC)*(-log10(Japanese_dge_Red_ref_SexMale_Myeloid_pDC[[i]]$adj.P.Val))
names(Japanese_dge_Red_ref_SexMale_Myeloid_pDC_rankings[[i]]) <- Japanese_dge_Red_ref_SexMale_Myeloid_pDC[[i]]$Gene
Japanese_dge_Red_ref_SexMale_Myeloid_pDC_rankings[[i]] <- sort(Japanese_dge_Red_ref_SexMale_Myeloid_pDC_rankings[[i]], decreasing = TRUE)
Japanese_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = Japanese_dge_Red_ref_SexMale_Myeloid_pDC_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
Japanese_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres_select[[i]] <- select(Japanese_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
Japanese_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres_select[[i]]$Author_Annotation <- i
}

#NK
Japanese_dge_Red_ref_SexMale_NK <- list()
Japanese_dge_Red_ref_SexMale_NK_rankings <- list()
Japanese_dge_Red_ref_SexMale_NK_GSEAres <- list()
Japanese_dge_Red_ref_SexMale_NK_GSEAres_select <- list()
for (i in NK_Author_Annotation_celltype) {
Japanese_dge_Red_ref_SexMale_NK[[i]] <- data.frame(JP_AIDA_NK_dream_modelFit_Red_ref_SexMale[[i]])
Japanese_dge_Red_ref_SexMale_NK_rankings[[i]] <- sign(Japanese_dge_Red_ref_SexMale_NK[[i]]$logFC)*(-log10(Japanese_dge_Red_ref_SexMale_NK[[i]]$adj.P.Val))
names(Japanese_dge_Red_ref_SexMale_NK_rankings[[i]]) <- Japanese_dge_Red_ref_SexMale_NK[[i]]$Gene
Japanese_dge_Red_ref_SexMale_NK_rankings[[i]] <- sort(Japanese_dge_Red_ref_SexMale_NK_rankings[[i]], decreasing = TRUE)
Japanese_dge_Red_ref_SexMale_NK_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = Japanese_dge_Red_ref_SexMale_NK_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
Japanese_dge_Red_ref_SexMale_NK_GSEAres_select[[i]] <- select(Japanese_dge_Red_ref_SexMale_NK_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
Japanese_dge_Red_ref_SexMale_NK_GSEAres_select[[i]]$Author_Annotation <- i
}

#T
Japanese_dge_Red_ref_SexMale_T <- list()
Japanese_dge_Red_ref_SexMale_T_rankings <- list()
Japanese_dge_Red_ref_SexMale_T_GSEAres <- list()
Japanese_dge_Red_ref_SexMale_T_GSEAres_select <- list()
for (i in T_Author_Annotation_celltype) {
Japanese_dge_Red_ref_SexMale_T[[i]] <- data.frame(JP_AIDA_T_dream_modelFit_Red_ref_SexMale[[i]])
Japanese_dge_Red_ref_SexMale_T_rankings[[i]] <- sign(Japanese_dge_Red_ref_SexMale_T[[i]]$logFC)*(-log10(Japanese_dge_Red_ref_SexMale_T[[i]]$adj.P.Val))
names(Japanese_dge_Red_ref_SexMale_T_rankings[[i]]) <- Japanese_dge_Red_ref_SexMale_T[[i]]$Gene
Japanese_dge_Red_ref_SexMale_T_rankings[[i]] <- sort(Japanese_dge_Red_ref_SexMale_T_rankings[[i]], decreasing = TRUE)
Japanese_dge_Red_ref_SexMale_T_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = Japanese_dge_Red_ref_SexMale_T_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
Japanese_dge_Red_ref_SexMale_T_GSEAres_select[[i]] <- select(Japanese_dge_Red_ref_SexMale_T_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
Japanese_dge_Red_ref_SexMale_T_GSEAres_select[[i]]$Author_Annotation <- i
}


#B
Japanese_dge_Red_ref_SexMale_B <- list()
Japanese_dge_Red_ref_SexMale_B_rankings <- list()
Japanese_dge_Red_ref_SexMale_B_GSEAres <- list()
Japanese_dge_Red_ref_SexMale_B_GSEAres_select <- list()
for (i in B_Author_Annotation_celltype) {
Japanese_dge_Red_ref_SexMale_B[[i]] <- data.frame(JP_AIDA_B_dream_modelFit_Red_ref_SexMale[[i]])
Japanese_dge_Red_ref_SexMale_B_rankings[[i]] <- sign(Japanese_dge_Red_ref_SexMale_B[[i]]$logFC)*(-log10(Japanese_dge_Red_ref_SexMale_B[[i]]$adj.P.Val))
names(Japanese_dge_Red_ref_SexMale_B_rankings[[i]]) <- Japanese_dge_Red_ref_SexMale_B[[i]]$Gene
Japanese_dge_Red_ref_SexMale_B_rankings[[i]] <- sort(Japanese_dge_Red_ref_SexMale_B_rankings[[i]], decreasing = TRUE)
Japanese_dge_Red_ref_SexMale_B_GSEAres[[i]] <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = Japanese_dge_Red_ref_SexMale_B_rankings[[i]],
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
Japanese_dge_Red_ref_SexMale_B_GSEAres_select[[i]] <- select(Japanese_dge_Red_ref_SexMale_B_GSEAres[[i]], pathway, padj, NES, leadingEdge) 
Japanese_dge_Red_ref_SexMale_B_GSEAres_select[[i]]$Author_Annotation <- i
}
```

```{r}
#Merge all cell types to get all GSEA results
Japanese_dge_Red_ref_SexMale_All_celltypes_GSEAres_df <- rbind(Japanese_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres_select[["Myeloid_unknown"]], Japanese_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres_select[["CD14+_Monocyte"]], Japanese_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres_select[["CD16+_Monocyte"]], Japanese_dge_Red_ref_SexMale_Myeloid_pDC_GSEAres_select[["cDC2"]], Japanese_dge_Red_ref_SexMale_NK_GSEAres_select[["NK_unknown"]],
Japanese_dge_Red_ref_SexMale_NK_GSEAres_select[["CD16+_NK"]], Japanese_dge_Red_ref_SexMale_NK_GSEAres_select[["CD56+_NK"]],
Japanese_dge_Red_ref_SexMale_T_GSEAres_select[["T_unknown"]], Japanese_dge_Red_ref_SexMale_T_GSEAres_select[["CD4+_T_unknown"]], Japanese_dge_Red_ref_SexMale_T_GSEAres_select[["CD4+_T_naive"]], Japanese_dge_Red_ref_SexMale_T_GSEAres_select[["CD4+_T_cm"]], Japanese_dge_Red_ref_SexMale_T_GSEAres_select[["CD4+_T_cyt"]], Japanese_dge_Red_ref_SexMale_T_GSEAres_select[["CD4+_T_em"]], Japanese_dge_Red_ref_SexMale_T_GSEAres_select[["CD8+_T_unknown"]], Japanese_dge_Red_ref_SexMale_T_GSEAres_select[["CD8+_T_naive"]], Japanese_dge_Red_ref_SexMale_T_GSEAres_select[["CD8+_T_GZMBhi"]], Japanese_dge_Red_ref_SexMale_T_GSEAres_select[["CD8+_T_GZMKhi"]], Japanese_dge_Red_ref_SexMale_T_GSEAres_select[["MAIT"]], Japanese_dge_Red_ref_SexMale_T_GSEAres_select[["gdT_GZMBhi"]], Japanese_dge_Red_ref_SexMale_T_GSEAres_select[["gdT_GZMKhi"]], Japanese_dge_Red_ref_SexMale_T_GSEAres_select[["Treg"]], Japanese_dge_Red_ref_SexMale_B_GSEAres_select[["B_unknown"]], Japanese_dge_Red_ref_SexMale_B_GSEAres_select[["naive_B"]], Japanese_dge_Red_ref_SexMale_B_GSEAres_select[["memory_B_IGHMlo"]], Japanese_dge_Red_ref_SexMale_B_GSEAres_select[["memory_B_IGHMhi"]])
```

```{r}
#to discuss only pathway
TH_dge_blue_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma <- subset(TH_dge_blue_ref_SexMale_All_celltypes_GSEAres_df, pathway %in% c("HALLMARK_INTERFERON_GAMMA_RESPONSE", "HALLMARK_TNFA_SIGNALING_VIA_NFKB"))
SG_Malay_dge_blue_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma <- subset(SG_Malay_dge_blue_ref_SexMale_All_celltypes_GSEAres_df, pathway %in% c("HALLMARK_INTERFERON_GAMMA_RESPONSE", "HALLMARK_TNFA_SIGNALING_VIA_NFKB"))
SG_Chinese_dge_blue_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma <- subset(SG_Chinese_dge_blue_ref_SexMale_All_celltypes_GSEAres_df, pathway %in% c("HALLMARK_INTERFERON_GAMMA_RESPONSE", "HALLMARK_TNFA_SIGNALING_VIA_NFKB"))
SG_Indian_dge_Orange_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma <- subset(SG_Indian_dge_Orange_ref_SexMale_All_celltypes_GSEAres_df, pathway %in% c("HALLMARK_INTERFERON_GAMMA_RESPONSE", "HALLMARK_TNFA_SIGNALING_VIA_NFKB"))
Korean_dge_Red_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma <- subset(Korean_dge_Red_ref_SexMale_All_celltypes_GSEAres_df, pathway %in% c("HALLMARK_INTERFERON_GAMMA_RESPONSE", "HALLMARK_TNFA_SIGNALING_VIA_NFKB"))
Japanese_dge_Red_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma <- subset(Japanese_dge_Red_ref_SexMale_All_celltypes_GSEAres_df, pathway %in% c("HALLMARK_INTERFERON_GAMMA_RESPONSE", "HALLMARK_TNFA_SIGNALING_VIA_NFKB"))

#Format
TH_dge_blue_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma$Ethncity_pathway <- paste("Thai_", TH_dge_blue_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma$pathway, sep = "")
SG_Malay_dge_blue_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma$Ethncity_pathway <- paste("SG_Malay_", SG_Malay_dge_blue_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma$pathway, sep = "")
SG_Chinese_dge_blue_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma$Ethncity_pathway <- paste("SG_Chinese_", SG_Chinese_dge_blue_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma$pathway, sep = "")
SG_Indian_dge_Orange_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma$Ethncity_pathway <- paste("SG_Indian_", SG_Indian_dge_Orange_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma$pathway, sep = "")
Korean_dge_Red_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma$Ethncity_pathway <- paste("Korean_", Korean_dge_Red_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma$pathway, sep = "")
Japanese_dge_Red_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma$Ethncity_pathway <- paste("Japanese_", Japanese_dge_Red_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma$pathway, sep = "")

All_AIDA_GSEA_TNF_gamma_combined <- rbind(SG_Indian_dge_Orange_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma,
                                          SG_Malay_dge_blue_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma,
                                          TH_dge_blue_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma,
                                          SG_Chinese_dge_blue_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma,
                                          Korean_dge_Red_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma,
                                          Japanese_dge_Red_ref_SexMale_All_celltypes_GSEAres_df_TNF_gamma
)
```

```{r fig.height=10, fig.width=12}
All_AIDA_GSEA_TNF_gamma_combined$Author_Annotation <- factor(All_AIDA_GSEA_TNF_gamma_combined$Author_Annotation, levels = Author_Annotation_celltype)
#Fig. 4F
All_AIDA_GSEA_TNF_gamma_combined_plot <- ggplot(All_AIDA_GSEA_TNF_gamma_combined) +
  geom_point(aes(x = Author_Annotation, y = Ethncity_pathway, color = NES, size=-log10(padj))) +
  theme_bw() +
  theme(axis.text.x = element_text(size=rel(1.15), angle = 90, vjust = 0.5, hjust=1),
        axis.title = element_text(size=rel(1.15))) +
  xlab("") +
  ylab("Pathway") +
  ggtitle("Pathway analysis by sex differences of each AIDA population") +
  theme(plot.title = element_text(hjust=0.5,	face = "bold"))  +
  scale_color_gradientn(name = "NES", colors = sex_col_gradient) +
  theme(legend.title = element_text(size=rel(1.15), hjust=0.5, face="bold")) + 
  facet_grid(pathway ~ ., scales='free') + 
  theme(text = element_text(size = 14,  color = "black"), strip.background = element_rect(fill = "#a1d76a", color = "black"))  
All_AIDA_GSEA_TNF_gamma_combined_plot
```

#8. Save images
```{r}
#Fig.4
pdf("/home/benjamaporn/AIDA/AIDA_Picture_manuscript/Part4_AIDA_MS/Fig.5A.pdf", width = 10, height = 4)
par(family = "Helvetica")
TH_DEGs_sig_Blue_ref_counts_plot
dev.off()

pdf("/home/benjamaporn/AIDA/AIDA_Picture_manuscript/Part4_AIDA_MS/Fig.5B.pdf", width = 12, height = 14)
par(family = "Helvetica")
TH_AIDA_Allcelltypes_dream_modelFit_Blue_ref_SexMale_df_representative_genes_mtx_plot
dev.off()

pdf("/home/benjamaporn/AIDA/AIDA_Picture_manuscript/Part4_AIDA_MS/Fig.5C.pdf", width = 10, height = 4)
par(family = "Helvetica")
TH_DEGs_sig_Blue_ref_SexMale_noYlinkedGenes_counts_plot
dev.off()

pdf("/home/benjamaporn/AIDA/AIDA_Picture_manuscript/Part4_AIDA_MS/Fig.5D.pdf", width = 9, height = 6)
par(family = "Helvetica")
TH_dge_blue_ref_SexMale_CD16NK_noYlinkedGenes_diffexpressed_newcutoff_plot
dev.off()

pdf("/home/benjamaporn/AIDA/AIDA_Picture_manuscript/Part4_AIDA_MS/Fig.S28.pdf", width = 12, height = 5.5)
par(family = "Helvetica")
bulk_ALL_AIDA_CD16NK_integrated_passQC_DCP_hvg_DGE_vobjGenes_mtx_df_merge_DAPK1_Sex_plot
dev.off()

pdf("/home/benjamaporn/AIDA/AIDA_Picture_manuscript/Part4_AIDA_MS/Fig.5E.pdf", width = 12, height = 10)
par(family = "Helvetica")
All_AIDA_GSEA_TNF_gamma_combined_plot
dev.off()
```