##Association Tests 
```{r}
library(plinkQC)
library(tidyverse)
library(ggplot2)
library(readr)
library(readxl)
library(dplyr)
library(stringr)
library(openxlsx)
library(genio)

#Input
chr6.dose <- read.table("/mnt/icbs_shared_storage_general_2022/Benjamaporn.sri/AIDA/HLA_2026/HLA/chr6.dose.raw", header=T)
chr6.dose_plink <- read_plink("/mnt/icbs_shared_storage_general_2022/Benjamaporn.sri/AIDA/HLA_2026/HLA/chr6.dose", verbose = TRUE)

#subset column that contain only HLA
chr6_markers_select <- subset(chr6.dose, select = -c(IID, FID, PAT, MAT, SEX, PHENOTYPE))
chr6_markers_colnames <- as.vector(colnames(chr6_markers_select))

#Read Files
#AIDA
Step03_Integration_integrated_PBMC_after_CellTypeSpecificQC_celltype_4level_July2024_AIDA_admix_ref_GSA_unrelated_samples <- readRDS("/mnt/icbs_shared_storage_general_2022/Benjamaporn/Analysis2025/Analysis202510/Step03_Integration_integrated_PBMC_after_CellTypeSpecificQC_celltype_4level_July2024_AIDA_admix_ref_GSA_unrelated_samples.rds")

Allcelltype_AIDA_admix_ref_GSA_unrelated_samples <- Step03_Integration_integrated_PBMC_after_CellTypeSpecificQC_celltype_4level_July2024_AIDA_admix_ref_GSA_unrelated_samples@meta.data 

#metadata
Allcelltype_AIDA_admix_ref_GSA_unrelated_samples_select <- select(Allcelltype_AIDA_admix_ref_GSA_unrelated_samples, Genotyping_ID, Age, Sex, Blue_ref, Orange_ref, Red_ref, Ethnicity, Country) %>%  
rename("IID" = "Genotyping_ID")  %>% 
distinct(IID, .keep_all = TRUE) 
rownames(Allcelltype_AIDA_admix_ref_GSA_unrelated_samples_select) <- NULL
AIDA_HLA_merged_metadata <- merge(chr6.dose, Allcelltype_AIDA_admix_ref_GSA_unrelated_samples_select, by = "IID")

#HLA and aa format
chr6.dose_plink_bim <- data.frame(chr6.dose_plink[["bim"]]) 
chr6.dose_plink_bim$Chr6_ID <- paste(chr6.dose_plink_bim$id, chr6.dose_plink_bim$alt, sep ="_") 
chr6.dose_plink_bim <- chr6.dose_plink_bim %>% mutate(across('Chr6_ID', str_replace, '[*]', '.')) %>% mutate(across('Chr6_ID', str_replace, ':', '.'))
row.names(chr6.dose_plink_bim) <- chr6.dose_plink_bim$Chr6_ID

#HLA markers
AIDA_HLA_select <- subset(AIDA_HLA_merged_metadata, select = c(IID, grep("HLA", names(AIDA_HLA_merged_metadata))))
row.names(AIDA_HLA_select) <- AIDA_HLA_select$IID
AIDA_HLA_select$IID <- NULL
AIDA_HLA_select_mtx <- as.matrix(AIDA_HLA_select)
```

#1. Association between Blue ancestry and HLA/amino acid polymorphisms alleles
```{r}
#Regression analysis
All_AIDA_HLA_Blue_ref <- list()
All_AIDA_HLA_Blue_ref_cof_list <- list()
All_AIDA_HLA_Blue_ref_cof_rename <- list()
All_AIDA_HLA_Blue_ref_cof_p <- list()
All_AIDA_HLA_Blue_ref_cof_p_rename <- list()
All_AIDA_HLA_Blue_ref_cof_p_onlyHLA <- list()

for (i in chr6_markers_colnames) {
All_AIDA_HLA_Blue_ref[[i]] <- lm(paste("Blue_ref ~", paste(c(i, "Age", "Sex"), collapse = " + ")), data = AIDA_HLA_merged_metadata)
All_AIDA_HLA_Blue_ref_cof_list[[i]] <- as.data.frame(All_AIDA_HLA_Blue_ref[[i]][["coefficients"]])
All_AIDA_HLA_Blue_ref_cof_rename[[i]] <- setNames(All_AIDA_HLA_Blue_ref_cof_list[[i]], i)
All_AIDA_HLA_Blue_ref_cof_p[[i]] <- as.data.frame(summary(All_AIDA_HLA_Blue_ref[[i]])$coefficients[,4])
All_AIDA_HLA_Blue_ref_cof_p_rename[[i]] <- setNames(All_AIDA_HLA_Blue_ref_cof_p[[i]], i)
All_AIDA_HLA_Blue_ref_cof_p_onlyHLA[[i]] <- All_AIDA_HLA_Blue_ref_cof_p_rename[[i]][rownames(All_AIDA_HLA_Blue_ref_cof_p_rename[[i]]) %in% i, ] 
}

#p-value
All_AIDA_HLA_Blue_ref_cof_p_onlyHLA_clean_list <- All_AIDA_HLA_Blue_ref_cof_p_onlyHLA[lengths(All_AIDA_HLA_Blue_ref_cof_p_onlyHLA) > 0]
All_AIDA_HLA_Blue_ref_cof_p_onlyHLA_clean_list_df <- data.frame(All_AIDA_HLA_Blue_ref_cof_p_onlyHLA_clean_list)
All_AIDA_HLA_Blue_ref_cof_p_onlyHLA_clean_list_mtx <- t(as.matrix(All_AIDA_HLA_Blue_ref_cof_p_onlyHLA_clean_list_df))

#combine p-value from HLA associaition results and chromosome position of each allele
All_AIDA_HLA_Blue_ref_p_chr6pos <- merge(chr6.dose_plink_bim, All_AIDA_HLA_Blue_ref_cof_p_onlyHLA_clean_list_mtx, by="row.names") %>% rename("P" = "V1")
All_AIDA_HLA_Blue_ref_p_chr6pos_hla <- subset(All_AIDA_HLA_Blue_ref_p_chr6pos, Chr6_ID %in% as.vector(colnames(AIDA_HLA_select)))

#only HLA region
All_AIDA_HLA_Blue_ref_p_chr6pos_plot <- ggplot(All_AIDA_HLA_Blue_ref_p_chr6pos, aes(x=pos, y=-log10(P))) +
    geom_point(aes(color=as.factor(chr)), alpha=0.8, size=2) +
    scale_color_manual(values = rep(c("grey", "#41b6c4"), 22 )) + 
    theme_bw() +
    theme( 
      legend.position="none",
      panel.border = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank()
    ) +
    ylim(0, 30) +
    geom_point(data=subset(All_AIDA_HLA_Blue_ref_p_chr6pos, P < 0.05), color="#41b6c4", size=1) +
    geom_label_repel(data=subset(All_AIDA_HLA_Blue_ref_p_chr6pos_hla, P < 1e-10), aes(label=Chr6_ID), size = 2, max.overlaps = Inf, fill = "white", box.padding = 0.5) + 
    labs(title = "MHC region at Chromosome 6: Association between HLA alleles and Blue ancestry across all AIDA populations (P < 1e-10)")
```

```{r}
rm(All_AIDA_HLA_Blue_ref)
rm(All_AIDA_HLA_Blue_ref_cof_list)
rm(All_AIDA_HLA_Blue_ref_cof_rename)
rm(All_AIDA_HLA_Blue_ref_cof_p)
rm(All_AIDA_HLA_Blue_ref_cof_p_rename)
rm(All_AIDA_HLA_Blue_ref_cof_p_onlyHLA)
```

#checked: subset only HLA allele in the Thai cohort
```{r}
#Regression analysis
TH_AIDA_HLA_Blue_ref <- list()
TH_AIDA_HLA_Blue_ref_cof_list <- list()
TH_AIDA_HLA_Blue_ref_cof_rename <- list()
TH_AIDA_HLA_Blue_ref_cof_p <- list()
TH_AIDA_HLA_Blue_ref_cof_p_rename <- list()
TH_AIDA_HLA_Blue_ref_cof_p_onlyHLA <- list()

for (i in chr6_markers_colnames) {
TH_AIDA_HLA_Blue_ref[[i]] <- lm(paste("Blue_ref ~", paste(c(i, "Age", "Sex"), collapse = " + ")), data = subset(AIDA_HLA_merged_metadata, Ethnicity == "Thai"))
TH_AIDA_HLA_Blue_ref_cof_list[[i]] <- as.data.frame(TH_AIDA_HLA_Blue_ref[[i]][["coefficients"]])
TH_AIDA_HLA_Blue_ref_cof_rename[[i]] <- setNames(TH_AIDA_HLA_Blue_ref_cof_list[[i]], i)
TH_AIDA_HLA_Blue_ref_cof_p[[i]] <- as.data.frame(summary(TH_AIDA_HLA_Blue_ref[[i]])$coefficients[,4])
TH_AIDA_HLA_Blue_ref_cof_p_rename[[i]] <- setNames(TH_AIDA_HLA_Blue_ref_cof_p[[i]], i)
TH_AIDA_HLA_Blue_ref_cof_p_onlyHLA[[i]] <- TH_AIDA_HLA_Blue_ref_cof_p_rename[[i]][rownames(TH_AIDA_HLA_Blue_ref_cof_p_rename[[i]]) %in% i, ] 
}

#p-value
TH_AIDA_HLA_Blue_ref_cof_p_onlyHLA_clean_list <- TH_AIDA_HLA_Blue_ref_cof_p_onlyHLA[lengths(TH_AIDA_HLA_Blue_ref_cof_p_onlyHLA) > 0]
TH_AIDA_HLA_Blue_ref_cof_p_onlyHLA_clean_list_df <- data.frame(TH_AIDA_HLA_Blue_ref_cof_p_onlyHLA_clean_list)
TH_AIDA_HLA_Blue_ref_cof_p_onlyHLA_clean_list_mtx <- t(as.matrix(TH_AIDA_HLA_Blue_ref_cof_p_onlyHLA_clean_list_df))

#combine p-value from HLA associaition results and chromosome position of each allele
TH_AIDA_HLA_Blue_ref_p_chr6pos <- merge(chr6.dose_plink_bim, TH_AIDA_HLA_Blue_ref_cof_p_onlyHLA_clean_list_mtx, by="row.names") %>% rename("P" = "V1")
TH_AIDA_HLA_Blue_ref_p_chr6pos_hla <- subset(TH_AIDA_HLA_Blue_ref_p_chr6pos, Chr6_ID %in% as.vector(colnames(AIDA_HLA_select)))

#only HLA region
ggplot(TH_AIDA_HLA_Blue_ref_p_chr6pos, aes(x=pos, y=-log10(P))) +
    geom_point( aes(color=as.factor(chr)), alpha=0.8, size=1) +
    scale_color_manual(values = rep(c("grey", "#41b6c4"), 22 )) + 
    theme_bw() +
    theme( 
      legend.position="none",
      panel.border = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank()
    ) +
    geom_label_repel(data=subset(TH_AIDA_HLA_Blue_ref_p_chr6pos_hla, P < 0.05), aes(label=Chr6_ID), size=3, max.overlaps = Inf) +
    geom_point(data=subset(TH_AIDA_HLA_Blue_ref_p_chr6pos, P < 0.05), color="#41b6c4", size=1) +
    labs(title = "MHC region at Chromosome 6: Association between HLA alleles and Blue ancestry (P < 0.05) in the Thai populations")
```

```{r}
rm(TH_AIDA_HLA_Blue_ref)
rm(TH_AIDA_HLA_Blue_ref_cof_list)
rm(TH_AIDA_HLA_Blue_ref_cof_rename)
rm(TH_AIDA_HLA_Blue_ref_cof_p)
rm(TH_AIDA_HLA_Blue_ref_cof_p_rename)
rm(TH_AIDA_HLA_Blue_ref_cof_p_onlyHLA)
```

#checked: subset only HLA allele in the SG cohort
```{r}
#Regression analysis
SG_AIDA_HLA_Blue_ref <- list()
SG_AIDA_HLA_Blue_ref_cof_list <- list()
SG_AIDA_HLA_Blue_ref_cof_rename <- list()
SG_AIDA_HLA_Blue_ref_cof_p <- list()
SG_AIDA_HLA_Blue_ref_cof_p_rename <- list()
SG_AIDA_HLA_Blue_ref_cof_p_onlyHLA <- list()

for (i in chr6_markers_colnames) {
SG_AIDA_HLA_Blue_ref[[i]] <- lm(paste("Blue_ref ~", paste(c(i, "Age", "Sex"), collapse = " + ")), data = subset(AIDA_HLA_merged_metadata, Country == "SG"))
SG_AIDA_HLA_Blue_ref_cof_list[[i]] <- as.data.frame(SG_AIDA_HLA_Blue_ref[[i]][["coefficients"]])
SG_AIDA_HLA_Blue_ref_cof_rename[[i]] <- setNames(SG_AIDA_HLA_Blue_ref_cof_list[[i]], i)
SG_AIDA_HLA_Blue_ref_cof_p[[i]] <- as.data.frame(summary(SG_AIDA_HLA_Blue_ref[[i]])$coefficients[,4])
SG_AIDA_HLA_Blue_ref_cof_p_rename[[i]] <- setNames(SG_AIDA_HLA_Blue_ref_cof_p[[i]], i)
SG_AIDA_HLA_Blue_ref_cof_p_onlyHLA[[i]] <- SG_AIDA_HLA_Blue_ref_cof_p_rename[[i]][rownames(SG_AIDA_HLA_Blue_ref_cof_p_rename[[i]]) %in% i, ] 
}

#p-value
SG_AIDA_HLA_Blue_ref_cof_p_onlyHLA_clean_list <- SG_AIDA_HLA_Blue_ref_cof_p_onlyHLA[lengths(SG_AIDA_HLA_Blue_ref_cof_p_onlyHLA) > 0]
SG_AIDA_HLA_Blue_ref_cof_p_onlyHLA_clean_list_df <- data.frame(SG_AIDA_HLA_Blue_ref_cof_p_onlyHLA_clean_list)
SG_AIDA_HLA_Blue_ref_cof_p_onlyHLA_clean_list_mtx <- t(as.matrix(SG_AIDA_HLA_Blue_ref_cof_p_onlyHLA_clean_list_df))

#combine p-value from HLA associaition results and chromosome position of each allele
SG_AIDA_HLA_Blue_ref_p_chr6pos <- merge(chr6.dose_plink_bim, SG_AIDA_HLA_Blue_ref_cof_p_onlyHLA_clean_list_mtx, by="row.names") %>% rename("P" = "V1")
SG_AIDA_HLA_Blue_ref_p_chr6pos_hla <- subset(SG_AIDA_HLA_Blue_ref_p_chr6pos, Chr6_ID %in% as.vector(colnames(AIDA_HLA_select)))

#only HLA region
ggplot(SG_AIDA_HLA_Blue_ref_p_chr6pos, aes(x=pos, y=-log10(P))) +
    geom_point( aes(color=as.factor(chr)), alpha=0.8, size=1) +
    scale_color_manual(values = rep(c("grey", "#41b6c4"), 22 )) + 
    theme_bw() +
    theme( 
      legend.position="none",
      panel.border = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank()
    ) +
    geom_label_repel(data=subset(SG_AIDA_HLA_Blue_ref_p_chr6pos_hla, P < 0.05), aes(label=Chr6_ID), size=2, max.overlaps = Inf) +
    geom_point(data=subset(SG_AIDA_HLA_Blue_ref_p_chr6pos, P < 0.05), color="#41b6c4", size=1) +
    labs(title = "MHC region at Chromosome 6: Association between HLA alleles and Blue ancestry (P < 0.05) in the SG populations")
```

```{r}
rm(SG_AIDA_HLA_Blue_ref)
rm(SG_AIDA_HLA_Blue_ref_cof_list)
rm(SG_AIDA_HLA_Blue_ref_cof_rename)
rm(SG_AIDA_HLA_Blue_ref_cof_p)
rm(SG_AIDA_HLA_Blue_ref_cof_p_rename)
rm(SG_AIDA_HLA_Blue_ref_cof_p_onlyHLA)
```

#Fig. 3G
```{r fig.height=4, fig.width=4}
library(ggpubr)
library(rstatix)
Allele_dosage_col_dict <- c(AA="#377EB8", AT="#4DAF4A", TT="#E41A1C")
AIDA_HLA_merged_metadata_HLA_DQB1.05.02_T <- select(AIDA_HLA_merged_metadata, IID, Blue_ref, HLA_DQB1.05.02_T, Country)
AIDA_HLA_merged_metadata_HLA_DQB1.05.02_T <- AIDA_HLA_merged_metadata_HLA_DQB1.05.02_T %>% 
                                            mutate(Allele_dosage = case_when(HLA_DQB1.05.02_T >= 0 & HLA_DQB1.05.02_T < 1 ~ "AA", 
                                                                             HLA_DQB1.05.02_T >= 1 & HLA_DQB1.05.02_T < 2 ~ "AT", 
                                                                             TRUE ~ "TT"))

#TH_AIDA
AIDA_HLA_merged_metadata_HLA_DQB1.05.02_T_dunn <- subset(AIDA_HLA_merged_metadata_HLA_DQB1.05.02_T, Country == "TH") %>%
  dunn_test(Blue_ref ~ Allele_dosage, p.adjust.method = "BH")
AIDA_HLA_merged_metadata_HLA_DQB1.05.02_T_dunn

AIDA_HLA_merged_metadata_HLA_DQB1.05.02_T_TH_plot <- ggboxplot(subset(AIDA_HLA_merged_metadata_HLA_DQB1.05.02_T, Country == "TH"), x="Allele_dosage", y="Blue_ref", outlier.shape=NA, size=0.15) +
geom_jitter(width = 0.08, size= 0.8, height=0, aes(color=Allele_dosage))+
ylab("Blue ancestry")+
xlab("")+
scale_color_manual(values=Allele_dosage_col_dict)+ 
guides(color = guide_legend(override.aes = list(size = 6))) +
labs(title = expression(italic("HLA_DQB1*05:02"))) +
stat_compare_means(label.x = 1.5, label.y = 1.05) + 
stat_pvalue_manual(AIDA_HLA_merged_metadata_HLA_DQB1.05.02_T_dunn, hide.ns = TRUE, y.position = 0.9)+ 
theme(text = element_text(size = 12), axis.text.x = element_blank(), legend.title = element_blank(), legend.position="right")
AIDA_HLA_merged_metadata_HLA_DQB1.05.02_T_TH_plot
```

#2. Association between Orange ancestry and HLA/amino acid polymorphisms alleles
```{r fig.height=5, fig.width=16}
#Regression analysis
All_AIDA_HLA_Orange_ref <- list()
All_AIDA_HLA_Orange_ref_cof_list <- list()
All_AIDA_HLA_Orange_ref_cof_rename <- list()
All_AIDA_HLA_Orange_ref_cof_p <- list()
All_AIDA_HLA_Orange_ref_cof_p_rename <- list()
All_AIDA_HLA_Orange_ref_cof_p_onlyHLA <- list()

for (i in chr6_markers_colnames) {
All_AIDA_HLA_Orange_ref[[i]] <- lm(paste("Orange_ref ~", paste(c(i, "Age", "Sex"), collapse = " + ")), data = AIDA_HLA_merged_metadata)
All_AIDA_HLA_Orange_ref_cof_list[[i]] <- as.data.frame(All_AIDA_HLA_Orange_ref[[i]][["coefficients"]])
All_AIDA_HLA_Orange_ref_cof_rename[[i]] <- setNames(All_AIDA_HLA_Orange_ref_cof_list[[i]], i)
All_AIDA_HLA_Orange_ref_cof_p[[i]] <- as.data.frame(summary(All_AIDA_HLA_Orange_ref[[i]])$coefficients[,4])
All_AIDA_HLA_Orange_ref_cof_p_rename[[i]] <- setNames(All_AIDA_HLA_Orange_ref_cof_p[[i]], i)
All_AIDA_HLA_Orange_ref_cof_p_onlyHLA[[i]] <- All_AIDA_HLA_Orange_ref_cof_p_rename[[i]][rownames(All_AIDA_HLA_Orange_ref_cof_p_rename[[i]]) %in% i, ] 
}

#p-value
All_AIDA_HLA_Orange_ref_cof_p_onlyHLA_clean_list <- All_AIDA_HLA_Orange_ref_cof_p_onlyHLA[lengths(All_AIDA_HLA_Orange_ref_cof_p_onlyHLA) > 0]
All_AIDA_HLA_Orange_ref_cof_p_onlyHLA_clean_list_df <- data.frame(All_AIDA_HLA_Orange_ref_cof_p_onlyHLA_clean_list)
All_AIDA_HLA_Orange_ref_cof_p_onlyHLA_clean_list_mtx <- t(as.matrix(All_AIDA_HLA_Orange_ref_cof_p_onlyHLA_clean_list_df))

#combine p-value from HLA associaition results and chromosome position of each allele
All_AIDA_HLA_Orange_ref_p_chr6pos <- merge(chr6.dose_plink_bim, All_AIDA_HLA_Orange_ref_cof_p_onlyHLA_clean_list_mtx, by="row.names") %>% rename("P" = "V1")
All_AIDA_HLA_Orange_ref_p_chr6pos_hla <- subset(All_AIDA_HLA_Orange_ref_p_chr6pos, Chr6_ID %in% as.vector(colnames(AIDA_HLA_select)))

#only HLA region
All_AIDA_HLA_Orange_ref_p_chr6pos_plot <- ggplot(All_AIDA_HLA_Orange_ref_p_chr6pos, aes(x=pos, y=-log10(P))) +
    geom_point( aes(color=as.factor(chr)), alpha=0.8, size=1.3) +
    scale_color_manual(values = rep(c("grey", "#fd8d3c"), 22 )) + 
    theme_bw() +
    theme( 
      legend.position="none",
      panel.border = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank()
    ) +
    ylim(0, 30) +
    geom_point(data=subset(All_AIDA_HLA_Orange_ref_p_chr6pos, P < 0.05), color="#fd8d3c", size=1) +
    geom_label_repel(data=subset(All_AIDA_HLA_Orange_ref_p_chr6pos_hla, P < 1e-10), aes(label=Chr6_ID), size=2, max.overlaps = Inf, fill = "white",  box.padding = 0.5) +
    labs(title = "MHC region at Chromosome 6: Association between HLA alleles and Orange ancestry across all AIDA populations (P < 1e-10)")
```

```{r}
rm(All_AIDA_HLA_Orange_ref)
rm(All_AIDA_HLA_Orange_ref_cof_list)
rm(All_AIDA_HLA_Orange_ref_cof_rename)
rm(All_AIDA_HLA_Orange_ref_cof_p)
rm(All_AIDA_HLA_Orange_ref_cof_p_rename)
rm(All_AIDA_HLA_Orange_ref_cof_p_onlyHLA)
```

#3. Association between Red ancestry and HLA/amino acid polymorphisms alleles
```{r fig.height=5, fig.width=16}
#Regression analysis
All_AIDA_HLA_Red_ref <- list()
All_AIDA_HLA_Red_ref_cof_list <- list()
All_AIDA_HLA_Red_ref_cof_rename <- list()
All_AIDA_HLA_Red_ref_cof_p <- list()
All_AIDA_HLA_Red_ref_cof_p_rename <- list()
All_AIDA_HLA_Red_ref_cof_p_onlyHLA <- list()

for (i in chr6_markers_colnames) {
All_AIDA_HLA_Red_ref[[i]] <- lm(paste("Red_ref ~", paste(c(i, "Age", "Sex"), collapse = " + ")), data = AIDA_HLA_merged_metadata)
All_AIDA_HLA_Red_ref_cof_list[[i]] <- as.data.frame(All_AIDA_HLA_Red_ref[[i]][["coefficients"]])
All_AIDA_HLA_Red_ref_cof_rename[[i]] <- setNames(All_AIDA_HLA_Red_ref_cof_list[[i]], i)
All_AIDA_HLA_Red_ref_cof_p[[i]] <- as.data.frame(summary(All_AIDA_HLA_Red_ref[[i]])$coefficients[,4])
All_AIDA_HLA_Red_ref_cof_p_rename[[i]] <- setNames(All_AIDA_HLA_Red_ref_cof_p[[i]], i)
All_AIDA_HLA_Red_ref_cof_p_onlyHLA[[i]] <- All_AIDA_HLA_Red_ref_cof_p_rename[[i]][rownames(All_AIDA_HLA_Red_ref_cof_p_rename[[i]]) %in% i, ] 
}

All_AIDA_HLA_Red_ref_cof_p_onlyHLA_clean_list <- All_AIDA_HLA_Red_ref_cof_p_onlyHLA[lengths(All_AIDA_HLA_Red_ref_cof_p_onlyHLA) > 0]
All_AIDA_HLA_Red_ref_cof_p_onlyHLA_clean_list_df <- data.frame(All_AIDA_HLA_Red_ref_cof_p_onlyHLA_clean_list)
All_AIDA_HLA_Red_ref_cof_p_onlyHLA_clean_list_mtx <- t(as.matrix(All_AIDA_HLA_Red_ref_cof_p_onlyHLA_clean_list_df))

#combine p-value from HLA associaition results and chromosome position of each allele
All_AIDA_HLA_Red_ref_p_chr6pos <- merge(chr6.dose_plink_bim, All_AIDA_HLA_Red_ref_cof_p_onlyHLA_clean_list_mtx, by="row.names") %>% rename("P" = "V1")
All_AIDA_HLA_Red_ref_p_chr6pos_hla <- subset(All_AIDA_HLA_Red_ref_p_chr6pos, Chr6_ID %in% as.vector(colnames(AIDA_HLA_select)))

#only HLA region
All_AIDA_HLA_Red_ref_p_chr6pos_plot <- ggplot(All_AIDA_HLA_Red_ref_p_chr6pos, aes(x=pos, y=-log10(P))) +
    geom_point( aes(color=as.factor(chr)), alpha=0.8, size=1.3) +
    scale_color_manual(values = rep(c("grey", "#bd0026"), 22 ))+ 
    theme_bw() +
    theme( 
      legend.position="none",
      panel.border = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank()
    ) +
    ylim(0, 30) +
    geom_point(data=subset(All_AIDA_HLA_Red_ref_p_chr6pos, P < 0.05), color="#bd0026", size=1) +
    geom_label_repel(data=subset(All_AIDA_HLA_Red_ref_p_chr6pos_hla, P < 1e-10), aes(label=Chr6_ID), size=2, max.overlaps = Inf, fill = "white", box.padding = 0.5) +
    labs(title = "MHC region at Chromosome 6: Association between HLA alleles and Red ancestry across all AIDA populations (P < 1e-10)")
```

```{r}
rm(All_AIDA_HLA_Red_ref)
rm(All_AIDA_HLA_Red_ref_cof_list)
rm(All_AIDA_HLA_Red_ref_cof_rename)
rm(All_AIDA_HLA_Red_ref_cof_p)
rm(All_AIDA_HLA_Red_ref_cof_p_rename)
rm(All_AIDA_HLA_Red_ref_cof_p_onlyHLA)
```

#Save main figure, supplementary figures and table
```{r}
#Fig. 3G
pdf("/home/benjamaporn.sri/HLA_analysis/Part3_HLA_MS/Fig.3G_HLA.pdf", width = 4.5, height = 4)
#Create a plot
par(family = "Helvetica")
AIDA_HLA_merged_metadata_HLA_DQB1.05.02_T_TH_plot
dev.off

#Supplementary figures
pdf("/home/benjamaporn.sri/HLA_analysis/Part3_HLA_MS/Fig.S20_All_AIDA_HLA_Blue_ref_p_chr6pos_plot.pdf", width = 16, height = 5)
#Create a plot
par(family = "Helvetica")
All_AIDA_HLA_Blue_ref_p_chr6pos_plot 
dev.off

pdf("/home/benjamaporn.sri/HLA_analysis/Part3_HLA_MS/Fig.S20_All_AIDA_HLA_Red_ref_p_chr6pos_plot.pdf", width = 16, height = 5)
#Create a plot
par(family = "Helvetica")
All_AIDA_HLA_Red_ref_p_chr6pos_plot 
dev.off

pdf("/home/benjamaporn.sri/HLA_analysis/Part3_HLA_MS/Fig.S20_All_AIDA_HLA_Orange_ref_p_chr6pos_plot.pdf", width = 16, height = 5)
#Create a plot
par(family = "Helvetica")
All_AIDA_HLA_Orange_ref_p_chr6pos_plot
dev.off


#Supplementary tables
All_AIDA_HLA_Blue_ref_p_chr6pos_select <- All_AIDA_HLA_Blue_ref_p_chr6pos  %>% select(chr, id, Chr6_ID, posg, pos, alt, ref, P) %>% rename("P (Blue ancestry)" = "P")
All_AIDA_HLA_Orange_ref_p_chr6pos_select <- All_AIDA_HLA_Orange_ref_p_chr6pos  %>% select(id, P) %>% rename("P (Orange ancestry)" = "P")
All_AIDA_HLA_Red_ref_p_chr6pos_select <- All_AIDA_HLA_Red_ref_p_chr6pos  %>% select(id, P) %>% rename("P (Red ancestry)" = "P")
All_AIDA_HLA_Ancestry_chr6pos_list <- list(All_AIDA_HLA_Blue_ref_p_chr6pos_select, All_AIDA_HLA_Orange_ref_p_chr6pos_select, All_AIDA_HLA_Red_ref_p_chr6pos_select)      
#merge all data frames together
All_AIDA_HLA_Ancestry_chr6pos_select <- All_AIDA_HLA_Ancestry_chr6pos_list %>% reduce(full_join, by='id')
write.xlsx(All_AIDA_HLA_Ancestry_chr6pos_select, file = "/home/benjamaporn.sri/HLA_analysis/Part3_HLA_MS/TableS10_All_AIDA_HLA_Ancestry_chr6pos_select.xlsx")
```

