#Load required library
```{r}
library(patchwork)
library(readr)
library(Seurat)
library(DESeq2)
library(tidyverse)
library(ggfortify)
library(ggrepel)
library(SeuratData)
library(SeuratWrappers)
library(Azimuth)
options(future.globals.maxSize = 100e9)
``` 

#1. Read processed data 
```{r}
#Read Files
#AIDA
Step03_Integration_integrated_PBMC_after_CellTypeSpecificQC_celltype_4level_July2024_AIDA_admix_ref_GSA_unrelated_samples <- readRDS("/mnt/icbs_shared_storage_general_2022/Benjamaporn/Analysis2025/Analysis202510/Step03_Integration_integrated_PBMC_after_CellTypeSpecificQC_celltype_4level_July2024_AIDA_admix_ref_GSA_unrelated_samples.rds")

#Only Myeloid innate cells
#Myeloid cells
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid <- readRDS("/home/benjamaporn/AIDA/GEX/AIDA_GEX/Celltypes/AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid.RDS")
#pDC
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_pDC <- readRDS("/home/benjamaporn/AIDA/GEX/AIDA_GEX/Celltypes/AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_pDC.RDS")
```
 
#2. Formatting seurat data
```{r}
#Metadata of unrelated Samples
AIDA_unrelated_samples_metadata <- Step03_Integration_integrated_PBMC_after_CellTypeSpecificQC_celltype_4level_July2024_AIDA_admix_ref_GSA_unrelated_samples@meta.data
rownames(AIDA_unrelated_samples_metadata) <- AIDA_unrelated_samples_metadata$barcode_name

#Myeloid cells
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_unrelated_samples <- subset(x = AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid, subset = barcode_name %in% AIDA_unrelated_samples_metadata$barcode_name) 
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_unrelated_samples_metadata <- AddMetaData(
  object = AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_unrelated_samples,
  metadata = AIDA_unrelated_samples_metadata
)

#pDC
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_pDC_unrelated_samples <- subset(x = AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_pDC, subset = barcode_name %in% AIDA_unrelated_samples_metadata$barcode_name) 
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_pDC_unrelated_samples_metadata <- AddMetaData(
  object = AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_pDC_unrelated_samples,
  metadata = AIDA_unrelated_samples_metadata
)

saveRDS(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_unrelated_samples_QC_samples_list, ".../AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_unrelated_samples_QC_samples_list.rds")
```

#3. Sample QC
```{r}
#No IN and sample with cell < 800 cells
DCP_ID_more800_ID <- AIDA_unrelated_samples_metadata %>%
  group_by(DCP_ID) %>% 
  summarise(n_cell = n_distinct(barcode_name)) %>% 
  filter(n_cell>800)
#checked
DCP_ID_less800_ID_related_LONZA_excluded <- subset(AIDA_unrelated_samples_metadata, !(DCP_ID %in% unique(DCP_ID_more800_ID$DCP_ID)))

#subset seurat object
#Myeloid cells
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_unrelated_samples_QC_samples <- subset(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_unrelated_samples_metadata, subset = DCP_ID %in% DCP_ID_more800_ID$DCP_ID)
#pDC
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_pDC_unrelated_samples_QC_samples <- subset(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_pDC_unrelated_samples_metadata, subset = DCP_ID %in% DCP_ID_more800_ID$DCP_ID)

AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_unrelated_samples_QC_samples_list <- list(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_unrelated_samples_QC_samples, AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_pDC_unrelated_samples_QC_samples)
```

#4. Fast integration using reciprocal PCA (RPCA): Myeloid + pDC 
```{r}
#Read file
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_unrelated_samples_QC_samples_list <- readRDS("/mnt/icbs_shared_storage_general_2022/Benjamaporn/Analysis2025/Analysis202510/AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_unrelated_samples_QC_samples_list.rds")
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged <- merge(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_unrelated_samples_QC_samples_list[[1]], AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_unrelated_samples_QC_samples_list[[2]])
#convert a v3 assay to a v5 assay: https://satijalab.org/seurat/articles/seurat5_integration
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged[["RNA"]] <- as(object = AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged[["RNA"]], Class = "Assay5")
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged[["RNA"]] <- split(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged[["RNA"]], f = AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged$Library)
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged
saveRDS(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged, ".../AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged")

#Follow https://satijalab.org/seurat/articles/seurat5_integration
#Read files
Output_PATH <- "/mnt/icbs_shared_storage_general_2022/Benjamaporn/Analysis2025/Analysis202510/myeloid_pDC_integration"
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged <-  readRDS(file.path(Output_PATH, "/AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged"))

#Before Integration
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged <- NormalizeData(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged)
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged <- FindVariableFeatures(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged)
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged <- ScaleData(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged)
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged <- RunPCA(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged)
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged <- FindNeighbors(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged, dims = 1:30, reduction = "pca")
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_unintegrated <- FindClusters(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged, resolution = 2, cluster.name = "unintegrated_clusters")
saveRDS(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_unintegrated, file.path(Output_PATH, "/AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_unintegrated.rds"))

#Integration
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_unintegrated <-  readRDS(file.path(Output_PATH, "/AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_unintegrated.rds"))
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_RPCA <- IntegrateLayers(object = AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_unintegrated, method = RPCAIntegration, orig.reduction = "pca", new.reduction = "integrated.rpca", verbose = FALSE)
saveRDS(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_RPCA, file.path(Output_PATH, "/AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_RPCA.rds"))

#For Visualization
#Before Integration
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_unintegrated <- readRDS(file.path(Output_PATH, "/AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_unintegrated.rds"))
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_unintegrated_clustered <- RunUMAP(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_unintegrated, dims = 1:30, reduction = "pca", reduction.name = "umap.unintegrated")
saveRDS(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_unintegrated_clustered, file.path(Output_PATH, "/AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_unintegrated_clustered.rds"))

#RPCA
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_RPCA <- readRDS(file.path(Output_PATH, "/AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_RPCA.rds"))
#Re-join layers after integration
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_RPCA[["RNA"]] <- JoinLayers(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_RPCA[["RNA"]])
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_RPCA <- FindNeighbors(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_RPCA, reduction = "integrated.rpca", dims = 1:30)
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_RPCA <- FindClusters(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_RPCA, resolution = 2, cluster.name = "rpca_clusters")
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_RPCA_joined_clustered <- RunUMAP(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_RPCA, reduction = "integrated.rpca", dims = 1:30, reduction.name = "umap.RPCA")
saveRDS(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_RPCA_joined_clustered, file.path(Output_PATH, "/AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_RPCA_joined_clustered.rds"))
```

#Differential Abundances by MiloR
#5. DA for Myeloid+pDC of only Thai populations
```{r}
#TH populations
library(miloR)
library(SingleCellExperiment)
library(scater)
library(scran)
library(dplyr)
library(patchwork)
library(readr)
library(Seurat)
library(dplyr)
library(scuttle)
library(irlba)
library(BiocParallel)
library(ggplot2)
library(sparseMatrixStats)

#Read AIDA file after merging Myeloid+pDC and then integration by rpca
Input_PATH <- "/mnt/icbs_shared_storage_general_2022/Benjamaporn/Analysis2025/Analysis202510/myeloid_pDC_integration"
AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_RPCA_joined_clustered <- readRDS(file.path(Input_PATH, "/AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_RPCA_joined_clustered.rds"))

#Subset only Thai AIDA
Output_PATH <- "/mnt/icbs_shared_storage_general_2022/Benjamaporn/Analysis2025/Analysis202511/DA_myeloid_pDC_integrated_rpca/Thai_DA"
#Subset only Thai
TH_Myeloid_pDC_integrated_RPCA <- subset(AIDA_Phase1_DataFreeze_v2_Step03_Seurat_JPKRSGTH_Myeloid_pDC_merged_RPCA_joined_clustered, subset = Ethnicity == "Thai")
saveRDS(TH_Myeloid_pDC_integrated_RPCA, file.path(Output_PATH, "/TH_Myeloid_pDC_integrated_RPCA.rds"))

#Create a Milo object
TH_Myeloid_pDC_integrated_RPCA_sce <- as.SingleCellExperiment(TH_Myeloid_pDC_integrated_RPCA)
TH_Myeloid_pDC_integrated_RPCA_milo <- Milo(TH_Myeloid_pDC_integrated_RPCA_sce)
saveRDS(TH_Myeloid_pDC_integrated_RPCA_milo, file.path(Output_PATH, "/TH_Myeloid_pDC_integrated_RPCA_milo.rds"))

##1_Preprocessing (at k=300, p=0.1))
##Applying with GLMM (per library) of Thai samples
##Construct KNN 
TH_Myeloid_pDC_integrated_RPCA_milo_buildGraph <- buildGraph(TH_Myeloid_pDC_integrated_RPCA_milo, k = 300, d = 30)

#2_Defining representative neighbourhoods
TH_Myeloid_pDC_integrated_RPCA_milo_make_nH <- makeNhoods(TH_Myeloid_pDC_integrated_RPCA_milo_buildGraph, prop = 0.1, k = 300, d = 30, refined = TRUE, refinement_scheme = "graph")

##3_Count Cells (per library)
colData(TH_Myeloid_pDC_integrated_RPCA_milo_make_nH)$ObsID <- paste(colData(TH_Myeloid_pDC_integrated_RPCA_milo_make_nH)$Library, colData(TH_Myeloid_pDC_integrated_RPCA_milo_make_nH)$DCP_ID, sep="_")
TH_Myeloid_pDC_integrated_RPCA_milo_count <- countCells(TH_Myeloid_pDC_integrated_RPCA_milo_make_nH, meta.data = data.frame(colData(TH_Myeloid_pDC_integrated_RPCA_milo_make_nH)), samples="ObsID")
saveRDS(TH_Myeloid_pDC_integrated_RPCA_milo_count, file.path(Output_PATH, "/TH_Myeloid_pDC_integrated_RPCA_milo_count.rds"))

#4_GLMM
TH_Myeloid_pDC_integrated_RPCA_milo_count <- readRDS(file.path(Output_PATH, "/TH_Myeloid_pDC_integrated_RPCA_milo_count.rds"))
#Metadata
TH_Myeloid_pDC_integrated_RPCA_milo_Obs <- data.frame(colData(TH_Myeloid_pDC_integrated_RPCA_milo_count))[,c("Library", "Genotyping_ID", "DCP_ID", "Ethnicity", "Age", "Sex", "Country.x", "Red_ref", "Blue_ref", "Orange_ref", "Red_GSA", "Blue_GSA", "Orange_GSA", "ObsID")]
TH_Myeloid_pDC_integrated_RPCA_milo_Obs <- distinct(TH_Myeloid_pDC_integrated_RPCA_milo_Obs)
rownames(TH_Myeloid_pDC_integrated_RPCA_milo_Obs) <- TH_Myeloid_pDC_integrated_RPCA_milo_Obs$ObsID
## Reorder rownames to match columns of nhoodCounts(milo)
TH_Myeloid_pDC_integrated_RPCA_milo_Obs <- TH_Myeloid_pDC_integrated_RPCA_milo_Obs[colnames(nhoodCounts(TH_Myeloid_pDC_integrated_RPCA_milo_count)), , drop=FALSE]
rownames(TH_Myeloid_pDC_integrated_RPCA_milo_Obs) <- TH_Myeloid_pDC_integrated_RPCA_milo_Obs$ObsID
saveRDS(TH_Myeloid_pDC_integrated_RPCA_milo_Obs, file.path(Output_PATH, "/TH_Myeloid_pDC_integrated_RPCA_milo_Obs.rds"))

#The impact of Blue ancestry
TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm <- testNhoods(TH_Myeloid_pDC_integrated_RPCA_milo_count, design = ~ Sex + Age + Blue_ref + (1|Library), design.df = TH_Myeloid_pDC_integrated_RPCA_milo_Obs, fdr.weighting = "graph-overlap", glmm.solver="Fisher")
saveRDS(TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm, file.path(Output_PATH, "/TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm.rds"))
```

#6. Visualization of Myeloid + pDC of Thai populations
```{r fig.height=5, fig.width=6.5}
DA_TH_Output_PATH <- "/mnt/icbs_shared_storage_general_2022/Benjamaporn/Analysis2025/Analysis202511/DA_myeloid_pDC_integrated_rpca/Thai_DA"
TH_Myeloid_pDC_integrated_RPCA <- readRDS(file.path(DA_TH_Output_PATH, "/TH_Myeloid_pDC_integrated_RPCA.rds"))
TH_Myeloid_pDC_integrated_RPCA_milo <- readRDS(file.path(DA_TH_Output_PATH, "/TH_Myeloid_pDC_integrated_RPCA_milo.rds"))
TH_Myeloid_pDC_integrated_RPCA_milo_count <- readRDS(file.path(DA_TH_Output_PATH, "/TH_Myeloid_pDC_integrated_RPCA_milo_count.rds"))
TH_Myeloid_pDC_integrated_RPCA_milo_Obs <- readRDS(file.path(DA_TH_Output_PATH, "/TH_Myeloid_pDC_integrated_RPCA_milo_Obs.rds"))

#GLMM results
TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm <- readRDS(file.path(DA_TH_Output_PATH, "/TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm.rds"))

#Fig. 3A (Top)
TH_Myeloid_pDC_integrated_RPCA@meta.data$Author_Annotation <- factor(TH_Myeloid_pDC_integrated_RPCA@meta.data$Author_Annotation, levels = c("Myeloid_unknown", "CD14+_Monocyte",  "CD16+_Monocyte",  "cDC1", "cDC2", "DC_SIGLEC6hi", "pDC"))
TH_Myeloid_pDC_integrated_RPCA_plot <- DimPlot(TH_Myeloid_pDC_integrated_RPCA, reduction = "umap.RPCA", group.by = "Author_Annotation", label = F, repel = TRUE, raster=FALSE, label.size = 3, cols = c("Myeloid_unknown"="#01579B", "CD16+_Monocyte"="#FFCA99",  "DC_SIGLEC6hi"="#006666", "cDC1"="#00FFFF", "cDC2"="#99FFFF", "CD14+_Monocyte"="#CC5500", "pDC"="#30BCAD")) + labs(title = "Myeloid innate cells across Thai populations", caption = "Author Annotation: Thai AIDA (n = 25,040 cells)") + xlab("UMAP1") + ylab("UMAP2")
TH_Myeloid_pDC_integrated_RPCA_plot

#Add log normalized count to Milo object
TH_Myeloid_pDC_integrated_RPCA_milo_count <- logNormCounts(TH_Myeloid_pDC_integrated_RPCA_milo_count)
#checked 
TH_Myeloid_pDC_integrated_RPCA_milo_count_checked_logcount <- data.frame(TH_Myeloid_pDC_integrated_RPCA_milo_count@assays@data@listData[["logcounts"]])
TH_Myeloid_pDC_integrated_RPCA_milo_count_checked_count <- data.frame(TH_Myeloid_pDC_integrated_RPCA_milo_count@assays@data@listData[["counts"]])
TH_Myeloid_pDC_integrated_RPCA_milo_count_size_factor <- 
data.frame(TH_Myeloid_pDC_integrated_RPCA_milo_count@colData@listData[["sizeFactor"]])

#Fig. 3A (Bottom)
TH_Myeloid_pDC_integrated_RPCA_milo_count_buildnh <- buildNhoodGraph(TH_Myeloid_pDC_integrated_RPCA_milo_count)
#GLMM
#The impact of blue ancestry
TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm_plot <- plotNhoodGraphDA(TH_Myeloid_pDC_integrated_RPCA_milo_count_buildnh, TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm, alpha=0.05, layout="UMAP.RPCA") +
scale_fill_gradient2(low = "#8d372c", high = "#2c828d", mid = 'white', midpoint = 0, name = "logFC")
TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm_plot

#Fig. 3B
Myeloid_pDC_Author_Annotation_celltype_DA_GLMM <- c("Myeloid_unknown", "CD14+_Monocyte", "CD16+_Monocyte", "cDC2", "pDC")
#Annotate cells
TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm_annotate_Author_Annotation <- annotateNhoods(TH_Myeloid_pDC_integrated_RPCA_milo_count_buildnh, TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm, coldata_col = "Author_Annotation")

#Remove nhoods that cannot pass DA testing
TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm_annotate_Author_Annotation_noNA <- na.omit(TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm_annotate_Author_Annotation)
#subset only CD14+ Monocyte nhoods for DEGs testing
TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm_annotate_Author_Annotation_noNA_CD14_Monocyte <- filter(TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm_annotate_Author_Annotation_noNA, Author_Annotation == "CD14+_Monocyte")
TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm_annotate_Author_Annotation_noNA_CD14_Monocyte$NhoodGroup <- TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm_annotate_Author_Annotation_noNA_CD14_Monocyte$Author_Annotation

#Plot DA for overall cell abundances
TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm_annotate_Author_Annotation_noNA$Author_Annotation <- factor(TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm_annotate_Author_Annotation_noNA$Author_Annotation, levels = Myeloid_pDC_Author_Annotation_celltype_DA_GLMM)

#Ploting
TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_plotDAbeeswarm <- plotDAbeeswarm(TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm_annotate_Author_Annotation_noNA, group.by = "Author_Annotation", alpha= 0.05) + 
xlab("") +
ggtitle("Thai populations (N = 59)")+
theme_light()+
theme(text = element_text(size = 16)) + 
annotate(geom="text", x=5, y=2, label="Enriched with\nBlue ancestry", color="#2c828d") + 
annotate(geom="text", x=5, y=-2, label="Depleted with\nBlue ancestry", color="#8d372c") + 
xlab("") +
scale_color_gradient2(low = "#8d372c", high = "#2c828d", mid = 'white', midpoint = 0, name = "logFC")
TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_plotDAbeeswarm
```

#7. Gene expression of nhoods associated with Blue ancestry
```{r fig.height=12, fig.width=10}
#Integrate objects has logcounts as input for gene expression
#Finding gene signatures for neighbourhoods
##Exclude zero counts genes
TH_Myeloid_pDC_integrated_RPCA_keep.rows <- rowSums(logcounts(TH_Myeloid_pDC_integrated_RPCA_milo_count_buildnh)) != 0
TH_Myeloid_pDC_integrated_RPCA_milo_count_buildnh <- TH_Myeloid_pDC_integrated_RPCA_milo_count_buildnh[TH_Myeloid_pDC_integrated_RPCA_keep.rows, ]

## Find HVGs
TH_Myeloid_pDC_integrated_RPCA_dec <- modelGeneVar(TH_Myeloid_pDC_integrated_RPCA_milo_count_buildnh)
TH_Myeloid_pDC_integrated_RPCA_hvgs <- getTopHVGs(TH_Myeloid_pDC_integrated_RPCA_dec, n=2000)

#The impact of Blue ancestry 
#DA results of only CD14+_Monocyte nhoods
TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm_annotate_Author_Annotation_noNA_CD14_Monocyte

#DGE testing within a nhood group 
#No need to perform findNhoodGroupMarkers() since those nhoods have already annotated
TH_dge_blue_ref_CD14_Monocyte  <- testDiffExp(TH_Myeloid_pDC_integrated_RPCA_milo_count_buildnh, TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm_annotate_Author_Annotation_noNA_CD14_Monocyte, design = ~ Sex + Age + Blue_ref, meta.data = data.frame(colData(TH_Myeloid_pDC_integrated_RPCA_milo_count_buildnh)), subset.row = TH_Myeloid_pDC_integrated_RPCA_hvgs)
TH_dge_blue_ref_CD14_Monocyte_df <- as.data.frame(TH_dge_blue_ref_CD14_Monocyte[["CD14+_Monocyte"]])

#Identifying signatures of DA subpopulations
#Ancestry
TH_dge_blue_ref_CD14_Monocyte_df$GeneID <- row.names(TH_dge_blue_ref_CD14_Monocyte_df)
TH_dge_blue_ref_CD14_Monocyte_gene_markers <- TH_dge_blue_ref_CD14_Monocyte_df[which(TH_dge_blue_ref_CD14_Monocyte_df$adj.P.Val < 0.05), "GeneID"]

#compute gene x nhood expression matrix
TH_CD14_Monocyte_integrated_RPCA_milo_count_buildnh_gex_blue_ref <- calcNhoodExpression(TH_Myeloid_pDC_integrated_RPCA_milo_count_buildnh, subset.row=TH_dge_blue_ref_CD14_Monocyte_gene_markers)

#Plotting
#only Upregulated genes with blue ancestry
TH_dge_blue_ref_CD14_Monocyte_df_UP <- filter(TH_dge_blue_ref_CD14_Monocyte_df, logFC > 0.5 & adj.P.Val < 0.05)
TH_CD14_Monocyte_integrated_RPCA_milo_count_buildnh_gex_blue_ref_heatmap_plot <- plotNhoodExpressionDA(TH_CD14_Monocyte_integrated_RPCA_milo_count_buildnh_gex_blue_ref, TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm_annotate_Author_Annotation_noNA_CD14_Monocyte, features = TH_dge_blue_ref_CD14_Monocyte_df_UP$GeneID,
                      assay="logcounts", alpha = 0.05,
                      scale_to_1 = TRUE, cluster_features = TRUE
                      )+
theme(text = element_text(size = 14))
TH_CD14_Monocyte_integrated_RPCA_milo_count_buildnh_gex_blue_ref_heatmap_plot
```

#8. Gene density of gene of interest
```{r fig.height=3, fig.width=4}
#Plot density
library("Nebulosa")

#Fig. 3D
#plot individual genes
#upregulated by blue ancestry
TH_Myeloid_pDC_integrated_RPCA_MXD1_density <- plot_density(TH_Myeloid_pDC_integrated_RPCA, "MXD1", pal = "viridis") + theme(plot.title = element_text(face = "italic")) + xlab("UMAP1") + ylab("UMAP2")
TH_Myeloid_pDC_integrated_RPCA_NFKBIA_density <- plot_density(TH_Myeloid_pDC_integrated_RPCA, "NFKBIA", pal = "viridis") + theme(plot.title = element_text(face = "italic")) + xlab("UMAP1") + ylab("UMAP2")
TH_Myeloid_pDC_integrated_RPCA_OSM <- plot_density(TH_Myeloid_pDC_integrated_RPCA, "OSM", pal = "viridis") + theme(plot.title = element_text(face = "italic")) + xlab("UMAP1") + ylab("UMAP2")
TH_Myeloid_pDC_integrated_RPCA_HLADQB1 <- plot_density(TH_Myeloid_pDC_integrated_RPCA, "HLA-DQB1", pal = "viridis") + theme(plot.title = element_text(face = "italic")) + xlab("UMAP1") + ylab("UMAP2")
TH_Myeloid_pDC_integrated_RPCA_MXD1_density
```

#9. Pathway analysis
```{r}
#Set seed
set.seed(123456)

library(tidyverse) 
library(RColorBrewer) 
library(fgsea)
library(msigdbr)
```

```{r}
#Retrieve the gene sets data frame 
human_gene_sets_df <- msigdbr(db_species = "HS", species = 'Homo sapiens', category = 'H')

# Convert to named list format required by fgsea
human_gene_sets_gsea <- human_gene_sets_df %>% split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(human_gene_sets_gsea), "gene sets\n"))

#Patways
human_gene_sets_gsea  <- human_gene_sets_gsea[!is.na(names(human_gene_sets_gsea))]
```

```{r fig.height=8, fig.width=8}
#Pathways from All signature genes of DA results of CD14+ Monocytes by Blue ancestry
TH_dge_blue_ref_CD14_Monocyte_df_rankings <- sign(TH_dge_blue_ref_CD14_Monocyte_df$logFC)*(-log10(TH_dge_blue_ref_CD14_Monocyte_df $adj.P.Val))
names(TH_dge_blue_ref_CD14_Monocyte_df_rankings) <- TH_dge_blue_ref_CD14_Monocyte_df$GeneID
TH_dge_blue_ref_CD14_Monocyte_df_rankings <- sort(TH_dge_blue_ref_CD14_Monocyte_df_rankings, decreasing = TRUE)

#Run GSEA 
TH_dge_blue_ref_CD14_Monocyte_df_GSEAres <- fgsea(pathways = human_gene_sets_gsea, 
                 stats = TH_dge_blue_ref_CD14_Monocyte_df_rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

#Plotting
TH_dge_blue_ref_CD14_Monocyte_df_GSEAres_plot <- ggplot(TH_dge_blue_ref_CD14_Monocyte_df_GSEAres) +
  geom_point(aes(x = NES, y = reorder(pathway, NES), color = -log10(padj), size=size)) +
  theme_bw() +
  theme(axis.text.x = element_text(size=rel(1.15)),
        axis.title = element_text(size=rel(1.15))) +
  xlab("Normalized Enrichment Score") +
  ylab("Pathway") +
  ggtitle("CD14+ Monocytes among Thai Populations (N=59)") +
  theme(plot.title = element_text(hjust=0.5,	face = "bold"))  +
  scale_color_gradientn(name = "Significance \n (-log10(padj))", colors = hcl.colors(50, "Temps", rev = T)) +
  theme(legend.title = element_text(size=rel(1.15),
	hjust=0.5, 
	face="bold"))+
  guides(size = guide_legend(order=1),
         color = guide_colourbar(order=2))
TH_dge_blue_ref_CD14_Monocyte_df_GSEAres_plot
```

#10. Save figures
```{r}
#Fig. 3A
pdf("/home/benjamaporn/AIDA/AIDA_Picture_manuscript/Part3_AIDA_MS/20260531/Fig.3A_UMAP.pdf", width = 6.5, height = 5)
par(family = "Helvetica")
TH_Myeloid_pDC_integrated_RPCA_plot
dev.off()

pdf("/home/benjamaporn/AIDA/AIDA_Picture_manuscript/Part3_AIDA_MS/20260531/Fig.3A_DA.pdf", width = 5, height = 3)
par(family = "Helvetica")
TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_glmm_plot
dev.off()

#Fig. 3B
pdf("/home/benjamaporn/AIDA/AIDA_Picture_manuscript/Part3_AIDA_MS/20260531/Fig.3B.pdf", width = 8, height = 5)
par(family = "Helvetica")
TH_Myeloid_pDC_integrated_RPCA_milo_count_da_results_Library_blue_ref_plotDAbeeswarm
dev.off()

#Fig. 3C
pdf("/home/benjamaporn/AIDA/AIDA_Picture_manuscript/Part3_AIDA_MS/20260531/Fig.3C.pdf", width = 10, height = 16)
par(family = "Helvetica")
TH_CD14_Monocyte_integrated_RPCA_milo_count_buildnh_gex_blue_ref_heatmap_plot
dev.off()

#DA:Ancestry
pdf("/home/benjamaporn/AIDA/AIDA_Picture_manuscript/Part3_AIDA_MS/20260531/Fig.3D.pdf", width = 8, height = 6)
# 2. Create a plot
par(family = "Helvetica")
TH_Myeloid_pDC_integrated_RPCA_MXD1_density + 
TH_Myeloid_pDC_integrated_RPCA_NFKBIA_density +
TH_Myeloid_pDC_integrated_RPCA_OSM +
TH_Myeloid_pDC_integrated_RPCA_HLADQB1 +
plot_layout(ncol = 2, nrow = 2)
dev.off()
```




